### 2021-10-14 BHRC_Prelim_Analyses.Rmd
Preliminary BHRC Analyses
Nothing super convincing so far :/

---
title: "BHRC Preliminary Analyses"
output: html_notebook
---

## For these analyses, I'll look at how early life trauma (Threat and Deprivation factors) affects three metrics of performance on the Attention Network Task (ANT): Alerting, Orienting, and Conflict.

### Load and organize data:
```{r}
sle_factors<-readRDS("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/BHRC/thdp_20212904.rds")
ant_scores<-read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/BHRC/ANT SCORES 100-1700ms RT cutoffs.csv")
demo_data<-read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/BHRC/BHRCS_w0_Arielle.csv")
proposal.data<-readRDS("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/BHRC/PhD_117BHRC_2021_10_02.rds")

colnames(sle_factors)[2]<-"SUBJECTID"
colnames(demo_data)[2]<-"SUBJECTID"
combined.data1<-merge(demo_data,ant_scores,by="SUBJECTID")

proposal.data.w0<-proposal.data[proposal.data$redcap_event_name=="wave0_arm_1",c("ident", setdiff(colnames(proposal.data),colnames(combined.data1)))]

combined.data2<-merge(proposal.data.w0,sle_factors,by="ident")
all.data<-merge(combined.data1,combined.data2,by="SUBJECTID")
```

### Remove test set:
```{r}
set.seed(2021)
test=sample(seq(1015),200,replace=FALSE)
test.data<-all.data[test,]
all.data<-all.data[-test,]
```

### View raw data
```{r}
ggplot(data=all.data,aes(threat,ALERT))+geom_point()
ggplot(data=all.data,aes(threat,ORIENT))+geom_point()
ggplot(data=all.data,aes(threat,CONFLICT))+geom_point()

ggplot(data=all.data,aes(depriv,ALERT))+geom_point()
ggplot(data=all.data,aes(depriv,ORIENT))+geom_point()
ggplot(data=all.data,aes(depriv,CONFLICT))+geom_point()
```

### Remove outliers (>3sd) on Alerting, Orienting, and Conflict scores
```{r}
ALERT.3sd <- sd(all.data[!is.na(all.data$ALERT),]$ALERT)*3
ORIENT.3sd <- sd(all.data[!is.na(all.data$ORIENT),]$ORIENT)*3
CONFLICT.3sd <- sd(all.data[!is.na(all.data$CONFLICT),]$CONFLICT)*3

ALERT.mean <- mean(all.data[!is.na(all.data$ALERT),]$ALERT)
ORIENT.mean <- mean(all.data[!is.na(all.data$ORIENT),]$ORIENT)
CONFLICT.mean <- mean(all.data[!is.na(all.data$CONFLICT),]$CONFLICT)

# These will be separate data structures retaining as many subjects as possible for each of the 3 performance metrics
all.data.ALERT <- all.data[(all.data$ALERT>(ALERT.mean-ALERT.3sd) & all.data$ALERT<(ALERT.mean+ALERT.3sd)),]
all.data.ORIENT <- all.data[(all.data$ORIENT>(ORIENT.mean-ORIENT.3sd) & all.data$ORIENT<(ORIENT.mean+ORIENT.3sd)),]
all.data.CONFLICT <- all.data[(all.data$CONFLICT>(CONFLICT.mean-CONFLICT.3sd) & all.data$CONFLICT<(CONFLICT.mean+CONFLICT.3sd)),]

all.data.complete.set <- all.data[(all.data$ALERT>(ALERT.mean-ALERT.3sd) & all.data$ALERT<(ALERT.mean+ALERT.3sd)) & (all.data$ORIENT>(ORIENT.mean-ORIENT.3sd) & all.data$ORIENT<(ORIENT.mean+ORIENT.3sd)) & (all.data$CONFLICT>(CONFLICT.mean-CONFLICT.3sd) & all.data$CONFLICT<(CONFLICT.mean+CONFLICT.3sd)),]
```

### Run linear models using Threat and Deprivation scores to predict ANT performance metrics
```{r}
summary(lm(ALERT~age+gender+threat+depriv,data=all.data.ALERT))
summary(lm(ORIENT~age+gender+threat+depriv,data=all.data.ORIENT))
summary(lm(CONFLICT~age+gender+threat+depriv,data=all.data.CONFLICT))

summary(lm(ALERT~age+gender+threat+depriv+age*threat+age*depriv+gender*threat+gender*depriv,data=all.data.ALERT)) # sig gender x threat
summary(lm(ORIENT~age+gender+threat+depriv+age*threat+age*depriv+gender*threat+gender*depriv,data=all.data.ORIENT))
summary(lm(CONFLICT~age+gender+threat+depriv+age*threat+age*depriv+gender*threat+gender*depriv,data=all.data.CONFLICT)) # close age x threat

summary(lm(ALERT~age+gender+threat+depriv+age*threat+age*depriv+gender*threat+gender*depriv+age*gender*threat+age*gender*depriv,data=all.data.ALERT))
summary(lm(ORIENT~age+gender+threat+depriv+age*threat+age*depriv+gender*threat+gender*depriv+age*gender*threat+age*gender*depriv,data=all.data.ORIENT)) 
summary(lm(CONFLICT~age+gender+threat+depriv+age*threat+age*depriv+gender*threat+gender*depriv+age*gender*threat+age*gender*depriv,data=all.data.CONFLICT))

summary(lm(ALERT~age+gender+depriv+depriv*age+depriv*gender,data=all.data.ALERT)) 
summary(lm(ORIENT~age+gender+depriv+depriv*age+depriv*gender,data=all.data.ORIENT))
summary(lm(CONFLICT~age+gender+depriv+depriv*age+depriv*gender,data=all.data.CONFLICT))

summary(lm(ALERT~age+gender+threat+threat*age+threat*gender,data=all.data.ALERT)) # Close on gender x threat
summary(lm(ORIENT~age+gender+threat+threat*age+threat*gender,data=all.data.ORIENT))
summary(lm(CONFLICT~age+gender+threat+threat*age+threat*gender,data=all.data.CONFLICT))
```

### Plotting
```{r}

ggplot(data=all.data.ALERT[!is.na(all.data.ALERT$gender),],aes(depriv,ALERT))+geom_point(size=1,colour=all.data.ALERT[!is.na(all.data.ALERT$gender),]$age)+geom_smooth(method=lm)+ facet_wrap(~ factor(gender,labels=c("Gender 1","Gender 2")))

ggplot(data=all.data.ALERT[!is.na(all.data.ALERT$gender),],aes(depriv,ALERT))+geom_point(size=1,colour=all.data.ALERT[!is.na(all.data.ALERT$gender),]$gender)+geom_smooth(method=lm)+ facet_grid(factor(all.data.ALERT[!is.na(all.data.ALERT$gender),]$gender,labels=c("Gender 1","Gender 2"))~factor(round(all.data.ALERT[!is.na(all.data.ALERT$gender),]$age)))

ggplot(data=all.data.ALERT[!is.na(all.data.ALERT$gender),],aes(threat,ALERT))+geom_point(size=1,colour=all.data.ALERT[!is.na(all.data.ALERT$gender),]$age)+geom_smooth(method=lm)+ facet_wrap(~ factor(gender,labels=c("Gender 1","Gender 2")))

ggplot(data=all.data.ALERT[!is.na(all.data.ALERT$gender),],aes(threat,ALERT))+geom_point(size=1,colour=all.data.ALERT[!is.na(all.data.ALERT$gender),]$gender)+geom_smooth(method=lm)+ facet_grid(factor(gender,labels=c("Gender 1","Gender 2"))~factor(round(age)))

```

### Stepwise linear regression
```{r}
model1<-lm(ALERT~age+gender,data=all.data.ALERT)
model2<-lm(ALERT~age+gender+depriv,data=all.data.ALERT)
model3<-lm(ALERT~age+gender+depriv+depriv*age+depriv*gender,data=all.data.ALERT)
model4<-lm(ALERT~age+gender+depriv+depriv*age+depriv*gender+age*gender*depriv,data=all.data.ALERT)
anova(model1,model2,test = "Chisq")
anova(model2,model3,test = "Chisq")
anova(model3,model4,test = "Chisq")

model1<-lm(ALERT~age+gender,data=all.data.ALERT)
model2<-lm(ALERT~age+gender+threat,data=all.data.ALERT)
model3<-lm(ALERT~age+gender+threat+threat*age+threat*gender,data=all.data.ALERT)
model4<-lm(ALERT~age+gender+threat+threat*age+threat*gender+age*gender*threat,data=all.data.ALERT)
model5<-lm(ALERT~age+gender+threat+threat*gender,data=all.data.ALERT)
anova(model1,model2,test = "Chisq")
anova(model2,model3,test = "Chisq")
anova(model3,model4,test = "Chisq")
anova(model2,model5,test = "Chisq") # close 0.057
```

### What if we split by symptoms?

### Sandbox
```{r}

summary(lm(threat~age+ALERT+ORIENT+CONFLICT,data=all.data.complete.set[all.data.complete.set$gender==1,]))
summary(lm(depriv~age+ALERT+ORIENT+CONFLICT,data=all.data.complete.set[all.data.complete.set$gender==1,]))

summary(lm(threat~age+ALERT+ORIENT+CONFLICT,data=all.data.complete.set[all.data.complete.set$gender==2,]))
summary(lm(depriv~age+ALERT+ORIENT+CONFLICT,data=all.data.complete.set[all.data.complete.set$gender==2,]))

summary(lm(threat~age+gender+ALERT+ORIENT+CONFLICT,data=all.data.complete.set))
summary(lm(depriv~age+gender+ALERT+ORIENT+CONFLICT,data=all.data.complete.set))

```