### 2021-10-28 ABCD_Prelim_Analyses.Rmd
Added preliminary PFN analysis

---
title: "ABCD_Prelim_Analyses"
output: html_document
---

# Load data
```{r cars}
var_exp <- read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/variance_exp_by_subj.csv")
nihtb<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/NihTB/abcd_tbss01.txt",header=TRUE)
nihtb.baseline<-nihtb[nihtb$eventname=="baseline_year_1_arm_1",]
merged.data1 <- merge(var_exp,nihtb.baseline,by="subjectkey")
lmt<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/Thomp/lmtp201.txt",header=TRUE)
lmt.baseline<-lmt[lmt$eventname=="baseline_year_1_arm_1",]
merged.data2 <- merge(merged.data1,lmt.baseline[, c("subjectkey", setdiff(colnames(lmt.baseline),colnames(merged.data1)))],by="subjectkey")
ravlt<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/Thomp/abcd_ps01.txt",header=TRUE)
ravlt.baseline<-ravlt[ravlt$eventname=="baseline_year_1_arm_1",]
merged.data3 <- merge(merged.data2,ravlt.baseline[, c("subjectkey", setdiff(colnames(ravlt.baseline),colnames(merged.data2)))],by="subjectkey")
trauma<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/Strife/abcd_ptsd01.txt",header=TRUE)
trauma.baseline<-trauma[trauma$eventname=="baseline_year_1_arm_1",]
combined.data <- merge(merged.data3,trauma.baseline[, c("subjectkey", setdiff(colnames(trauma.baseline),colnames(merged.data3)))],by="subjectkey")
```

# Coerce data types
```{r}
combined.data$nihtbx_flanker_uncorrected <- as.numeric(combined.data$nihtbx_flanker_uncorrected)
combined.data$nihtbx_picvocab_uncorrected <- as.numeric(combined.data$nihtbx_picvocab_uncorrected)
combined.data$nihtbx_list_uncorrected <- as.numeric(combined.data$nihtbx_list_uncorrected)
combined.data$nihtbx_pattern_uncorrected <- as.numeric(combined.data$nihtbx_pattern_uncorrected)
combined.data$nihtbx_cardsort_uncorrected <- as.numeric(combined.data$nihtbx_cardsort_uncorrected)
combined.data$nihtbx_picture_uncorrected <- as.numeric(combined.data$nihtbx_picture_uncorrected)
combined.data$nihtbx_reading_uncorrected <- as.numeric(combined.data$nihtbx_reading_uncorrected)
combined.data$lmt_scr_perc_correct <- as.numeric(combined.data$lmt_scr_perc_correct)
combined.data$pea_ravlt_sd_trial_i_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_i_tc)
combined.data$pea_ravlt_sd_trial_ii_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_ii_tc)
combined.data$pea_ravlt_sd_trial_iii_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_iii_tc)
combined.data$pea_ravlt_sd_trial_iv_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_iv_tc)
combined.data$pea_ravlt_sd_trial_v_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_v_tc)
combined.data$interview_age <- as.numeric(combined.data$interview_age)

ind_pea_ravlt = c(which(names(combined.data)=="pea_ravlt_sd_trial_i_tc"),which(names(combined.data)=="pea_ravlt_sd_trial_ii_tc"),
	which(names(combined.data)=="pea_ravlt_sd_trial_iii_tc"),which(names(combined.data)=="pea_ravlt_sd_trial_iv_tc"),
	which(names(combined.data)=="pea_ravlt_sd_trial_v_tc")); names(combined.data)[ind_pea_ravlt]; summary(combined.data[,ind_pea_ravlt])
combined.data$pea_ravlt_ld = apply(combined.data[,ind_pea_ravlt],1,sum)

combined.data$pea_ravlt_ld <- as.numeric(combined.data$pea_ravlt_ld)

# Rename according to Thompson 2018 code
names(combined.data)[which(names(combined.data)=="nihtbx_picvocab_uncorrected")] = "PicVocab"
names(combined.data)[which(names(combined.data)=="nihtbx_flanker_uncorrected")] = "Flanker"
names(combined.data)[which(names(combined.data)=="nihtbx_list_uncorrected")] = "List"
names(combined.data)[which(names(combined.data)=="nihtbx_cardsort_uncorrected")] = "CardSort"
names(combined.data)[which(names(combined.data)=="nihtbx_pattern_uncorrected")] = "Pattern"
names(combined.data)[which(names(combined.data)=="nihtbx_picture_uncorrected")] = "Picture"
names(combined.data)[which(names(combined.data)=="nihtbx_reading_uncorrected")] = "Reading"
names(combined.data)[which(names(combined.data)=="pea_ravlt_ld")] = "RAVLT"
names(combined.data)[which(names(combined.data)=="lmt_scr_perc_correct")] = "LMT"
```

# Inclusion/Exclusion
```{r}
# First, to harmonize our criteria with UMinn, retain participants with at least 8min of clean data instead of 12min. TR = 0.8s so 8min = 480s = 600TRs
num_TRs <- read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/num_TRs_by_subj_redo.csv")
data<-merge(combined.data,num_TRs,by="subjectkey")
data.clean<-data[data$numTRs>=600,]

# Second, remove participants based on booleans from ABCD
abcd_imgincl01 <- read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/abcd_imgincl01.csv")
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$eventnam=="baseline_year_1_arm_1",]
abcd_imgincl01 <- abcd_imgincl01[!abcd_imgincl01$visit=="",]
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$imgincl_t1w_include==1,]
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$imgincl_rsfmri_include==1,]
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$imgincl_mid_include==1,]
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$imgincl_nback_include==1,]
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$imgincl_sst_include==1,]

combined.data <- merge(data.clean,abcd_imgincl01[, c("subjectkey",'imgincl_t1w_include')],by="subjectkey")
```

# Calculate Thompson factor scores
```{r}
combined.data$thompson_PC1 <- (0.754*combined.data$PicVocab)+(0.213*combined.data$Flanker)+(0.471*combined.data$List)+(0.205*combined.data$CardSort)+(0.015*combined.data$Pattern)+(0.012*combined.data$Picture)+(0.82*combined.data$Reading)+(0.306*combined.data$RAVLT)+(0.5*combined.data$LMT)

combined.data$thompson_PC2 <- (0.065*combined.data$PicVocab)+(0.712*combined.data$Flanker)+(0.148*combined.data$List)+(0.71*combined.data$CardSort)+(0.813*combined.data$Pattern)+(0.135*combined.data$Picture)+(0.12*combined.data$Reading)+(0.125*combined.data$RAVLT)+(0.299*combined.data$LMT)

combined.data$thompson_PC3 <- (0.19*combined.data$PicVocab)+(0.067*combined.data$Flanker)+(0.493*combined.data$List)+(0.232*combined.data$CardSort)+(0.085*combined.data$Pattern)+(0.863*combined.data$Picture)+(0.122*combined.data$Reading)+(0.712*combined.data$RAVLT)+(0.068*combined.data$LMT)
```

# Compare variance explained to Thompson factor scores
```{r}
summary(lm(varexp~interview_age+sex+thompson_PC1+thompson_PC2+thompson_PC3,data=combined.data))
model1<-lm(thompson_PC2~interview_age+sex,data=combined.data)
model2<-lm(thompson_PC2~interview_age+sex+varexp,data=combined.data)
anova(model1,model2,test = "Chisq")
```

# Compare variance explained to Trauma types
```{r}
combined.data[,225:241]<-as.numeric(unlist(combined.data[,225:241]))
combined.data$ACEsTotal <- rowSums(combined.data[,c("ksads_ptsd_raw_754_p","ksads_ptsd_raw_755_p","ksads_ptsd_raw_756_p","ksads_ptsd_raw_757_p","ksads_ptsd_raw_758_p","ksads_ptsd_raw_759_p","ksads_ptsd_raw_760_p","ksads_ptsd_raw_761_p","ksads_ptsd_raw_762_p","ksads_ptsd_raw_763_p","ksads_ptsd_raw_764_p","ksads_ptsd_raw_765_p","ksads_ptsd_raw_766_p","ksads_ptsd_raw_767_p","ksads_ptsd_raw_768_p","ksads_ptsd_raw_769_p","ksads_ptsd_raw_770_p")])

summary(lm(varexp~interview_age+sex+ksads_ptsd_raw_754_p+ksads_ptsd_raw_755_p+ksads_ptsd_raw_756_p+ksads_ptsd_raw_757_p+ksads_ptsd_raw_758_p+ksads_ptsd_raw_759_p+ksads_ptsd_raw_760_p+ksads_ptsd_raw_761_p+ksads_ptsd_raw_762_p+ksads_ptsd_raw_763_p+ksads_ptsd_raw_764_p+ksads_ptsd_raw_765_p+ksads_ptsd_raw_766_p+ksads_ptsd_raw_767_p+ksads_ptsd_raw_768_p+ksads_ptsd_raw_769_p+ksads_ptsd_raw_770_p,data=combined.data))

# Trauma endorsements are sparse (yay) so will use lasso
library(glmnet)
set.seed(1)
combined.data<-combined.data[complete.cases(combined.data$ACEsTotal),]
combined.data<-combined.data[complete.cases(combined.data$thompson_PC2),]
train=sample(seq(length(combined.data$subjectkey)),round((2/3)*length(combined.data$subjectkey)),replace=FALSE)
lasso.tr=glmnet(as.matrix(combined.data[train,c("ksads_ptsd_raw_754_p","ksads_ptsd_raw_755_p","ksads_ptsd_raw_756_p","ksads_ptsd_raw_757_p","ksads_ptsd_raw_758_p","ksads_ptsd_raw_759_p","ksads_ptsd_raw_760_p","ksads_ptsd_raw_761_p","ksads_ptsd_raw_762_p","ksads_ptsd_raw_763_p","ksads_ptsd_raw_764_p","ksads_ptsd_raw_765_p","ksads_ptsd_raw_766_p","ksads_ptsd_raw_767_p","ksads_ptsd_raw_768_p","ksads_ptsd_raw_769_p","ksads_ptsd_raw_770_p")]),combined.data[train,"varexp"],family="gaussian")
#lasso.tr
pred=predict(lasso.tr,as.matrix(combined.data[-train,c("ksads_ptsd_raw_754_p","ksads_ptsd_raw_755_p","ksads_ptsd_raw_756_p","ksads_ptsd_raw_757_p","ksads_ptsd_raw_758_p","ksads_ptsd_raw_759_p","ksads_ptsd_raw_760_p","ksads_ptsd_raw_761_p","ksads_ptsd_raw_762_p","ksads_ptsd_raw_763_p","ksads_ptsd_raw_764_p","ksads_ptsd_raw_765_p","ksads_ptsd_raw_766_p","ksads_ptsd_raw_767_p","ksads_ptsd_raw_768_p","ksads_ptsd_raw_769_p","ksads_ptsd_raw_770_p")]))
dim(pred)
rmse=sqrt(apply((combined.data[-train,"varexp"]-pred)^2,2,mean))
plot(log(lasso.tr$lambda),rmse,type="b",xlab="Log(lambda")
lam.best=lasso.tr$lambda[order(rmse)[1]] # extracting the best lambda
lam.best
coef(lasso.tr,s=lam.best)

summary(lm(varexp~interview_age+sex+ksads_ptsd_raw_755_p+ksads_ptsd_raw_756_p+ksads_ptsd_raw_757_p+ksads_ptsd_raw_758_p+ksads_ptsd_raw_762_p+ksads_ptsd_raw_763_p+ksads_ptsd_raw_764_p+ksads_ptsd_raw_766_p,data=combined.data))

model1<-lm(varexp~interview_age+sex,data=combined.data)
model2<-lm(varexp~interview_age+sex+ksads_ptsd_raw_755_p+ksads_ptsd_raw_756_p+ksads_ptsd_raw_757_p+ksads_ptsd_raw_758_p+ksads_ptsd_raw_762_p+ksads_ptsd_raw_763_p+ksads_ptsd_raw_764_p+ksads_ptsd_raw_766_p,data=combined.data)
anova(model1,model2,test = "Chisq")

```

```{r}

model1<-lm(varexp~interview_age+sex,data=combined.data)
model2<-lm(varexp~interview_age+sex+thompson_PC2,data=combined.data)
model3<-lm(varexp~interview_age+sex+thompson_PC2+interview_age*thompson_PC2,data=combined.data)
anova(model1,model2,test = "Chisq")
anova(model2,model3,test = "Chisq")

model1<-lm(thompson_PC2~interview_age+sex,data=combined.data)
model2<-lm(thompson_PC2~interview_age+sex+varexp,data=combined.data)
model3<-lm(thompson_PC2~interview_age+sex+varexp+interview_age*varexp,data=combined.data)
anova(model1,model2,test = "Chisq")
anova(model2,model3,test = "Chisq")

summary(lm(thompson_PC2~interview_age*sex*varexp,data=combined.data))
```

## Check for outliers (>3sd) on Thompson factor scores
```{r}
mean_PC1 <- mean(combined.data$thompson_PC1)
mean_PC2 <- mean(combined.data$thompson_PC2)
mean_PC3 <- mean(combined.data$thompson_PC3)
sd3_PC1 <- sd(combined.data$thompson_PC1)*3
sd3_PC2 <- sd(combined.data$thompson_PC2)*3
sd3_PC3 <- sd(combined.data$thompson_PC3)*3

sum(combined.data$thompson_PC1 > mean_PC1+sd3_PC1) # 3 outliers
sum(combined.data$thompson_PC2 > mean_PC2+sd3_PC2) # 0 outliers
sum(combined.data$thompson_PC3 > mean_PC3+sd3_PC3) # 2 outliers

combined.data.noOutliers <- combined.data[(combined.data$thompson_PC1>(mean_PC1-sd3_PC1) & combined.data$thompson_PC1<(mean_PC1+sd3_PC1) & combined.data$thompson_PC2>(mean_PC2-sd3_PC2) & combined.data$thompson_PC2<(mean_PC2+sd3_PC2) & combined.data$thompson_PC3>(mean_PC3-sd3_PC3) & combined.data$thompson_PC3<(mean_PC3+sd3_PC3)),]

summary(lm(thompson_PC2~interview_age*sex*varexp,data=combined.data.noOutliers))

```

## PFNs analysis (CLEAR BEFORE RUNNING)
# First load in the spatial extent of each PFN (they're in order, 1-17)
# Then merge with the same datasets as above (NIH toolbox, trauma, etc.)
```{r}
pfn_sizes <- read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/All_PFN_sizes.csv")
nihtb<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/NihTB/abcd_tbss01.txt",header=TRUE)
nihtb.baseline<-nihtb[nihtb$eventname=="baseline_year_1_arm_1",]
merged.data1 <- merge(pfn_sizes,nihtb.baseline,by="subjectkey")
lmt<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/Thomp/lmtp201.txt",header=TRUE)
lmt.baseline<-lmt[lmt$eventname=="baseline_year_1_arm_1",]
merged.data2 <- merge(merged.data1,lmt.baseline[, c("subjectkey", setdiff(colnames(lmt.baseline),colnames(merged.data1)))],by="subjectkey")
ravlt<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/Thomp/abcd_ps01.txt",header=TRUE)
ravlt.baseline<-ravlt[ravlt$eventname=="baseline_year_1_arm_1",]
merged.data3 <- merge(merged.data2,ravlt.baseline[, c("subjectkey", setdiff(colnames(ravlt.baseline),colnames(merged.data2)))],by="subjectkey")
trauma<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/Strife/abcd_ptsd01.txt",header=TRUE)
trauma.baseline<-trauma[trauma$eventname=="baseline_year_1_arm_1",]
combined.data <- merge(merged.data3,trauma.baseline[, c("subjectkey", setdiff(colnames(trauma.baseline),colnames(merged.data3)))],by="subjectkey")

# Coerce data types
combined.data$nihtbx_flanker_uncorrected <- as.numeric(combined.data$nihtbx_flanker_uncorrected)
combined.data$nihtbx_picvocab_uncorrected <- as.numeric(combined.data$nihtbx_picvocab_uncorrected)
combined.data$nihtbx_list_uncorrected <- as.numeric(combined.data$nihtbx_list_uncorrected)
combined.data$nihtbx_pattern_uncorrected <- as.numeric(combined.data$nihtbx_pattern_uncorrected)
combined.data$nihtbx_cardsort_uncorrected <- as.numeric(combined.data$nihtbx_cardsort_uncorrected)
combined.data$nihtbx_picture_uncorrected <- as.numeric(combined.data$nihtbx_picture_uncorrected)
combined.data$nihtbx_reading_uncorrected <- as.numeric(combined.data$nihtbx_reading_uncorrected)
combined.data$lmt_scr_perc_correct <- as.numeric(combined.data$lmt_scr_perc_correct)
combined.data$pea_ravlt_sd_trial_i_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_i_tc)
combined.data$pea_ravlt_sd_trial_ii_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_ii_tc)
combined.data$pea_ravlt_sd_trial_iii_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_iii_tc)
combined.data$pea_ravlt_sd_trial_iv_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_iv_tc)
combined.data$pea_ravlt_sd_trial_v_tc <- as.numeric(combined.data$pea_ravlt_sd_trial_v_tc)
combined.data$interview_age <- as.numeric(combined.data$interview_age)

ind_pea_ravlt = c(which(names(combined.data)=="pea_ravlt_sd_trial_i_tc"),which(names(combined.data)=="pea_ravlt_sd_trial_ii_tc"),
	which(names(combined.data)=="pea_ravlt_sd_trial_iii_tc"),which(names(combined.data)=="pea_ravlt_sd_trial_iv_tc"),
	which(names(combined.data)=="pea_ravlt_sd_trial_v_tc")); names(combined.data)[ind_pea_ravlt]; summary(combined.data[,ind_pea_ravlt])
combined.data$pea_ravlt_ld = apply(combined.data[,ind_pea_ravlt],1,sum)

combined.data$pea_ravlt_ld <- as.numeric(combined.data$pea_ravlt_ld)

# Rename according to Thompson 2018 code
names(combined.data)[which(names(combined.data)=="nihtbx_picvocab_uncorrected")] = "PicVocab"
names(combined.data)[which(names(combined.data)=="nihtbx_flanker_uncorrected")] = "Flanker"
names(combined.data)[which(names(combined.data)=="nihtbx_list_uncorrected")] = "List"
names(combined.data)[which(names(combined.data)=="nihtbx_cardsort_uncorrected")] = "CardSort"
names(combined.data)[which(names(combined.data)=="nihtbx_pattern_uncorrected")] = "Pattern"
names(combined.data)[which(names(combined.data)=="nihtbx_picture_uncorrected")] = "Picture"
names(combined.data)[which(names(combined.data)=="nihtbx_reading_uncorrected")] = "Reading"
names(combined.data)[which(names(combined.data)=="pea_ravlt_ld")] = "RAVLT"
names(combined.data)[which(names(combined.data)=="lmt_scr_perc_correct")] = "LMT"

num_TRs <- read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/num_TRs_by_subj_redo.csv")
data<-merge(combined.data,num_TRs,by="subjectkey")
data.clean<-data[data$numTRs>=600,] # we go from 9,132 to 8,293 at this step

# Remove participants based on booleans from ABCD
abcd_imgincl01 <- read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/abcd_imgincl01.csv")
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$eventnam=="baseline_year_1_arm_1",]
abcd_imgincl01 <- abcd_imgincl01[!abcd_imgincl01$visit=="",] # start here with 11,663 (8,287 of the 9,132 set)
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$imgincl_t1w_include==1,] # down to 11,124 (8,045 of the 9,132 set)
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$imgincl_rsfmri_include==1,] # down to 9,388 (7,475 of the 9,132 set)
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$imgincl_mid_include==1,] # down to 7,847 (6,344 of the 9,132 set)
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$imgincl_nback_include==1,] # down to 6,348 (5,161 of the 9,132 set)
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$imgincl_sst_include==1,] # down to 5,606 (4,164 of the 9,132 set)

combined.data <- merge(data.clean,abcd_imgincl01[, c("subjectkey",'imgincl_t1w_include')],by="subjectkey")

# Calculate Thompson factor scores
combined.data$thompson_PC1 <- (0.754*combined.data$PicVocab)+(0.213*combined.data$Flanker)+(0.471*combined.data$List)+(0.205*combined.data$CardSort)+(0.015*combined.data$Pattern)+(0.012*combined.data$Picture)+(0.82*combined.data$Reading)+(0.306*combined.data$RAVLT)+(0.5*combined.data$LMT)

combined.data$thompson_PC2 <- (0.065*combined.data$PicVocab)+(0.712*combined.data$Flanker)+(0.148*combined.data$List)+(0.71*combined.data$CardSort)+(0.813*combined.data$Pattern)+(0.135*combined.data$Picture)+(0.12*combined.data$Reading)+(0.125*combined.data$RAVLT)+(0.299*combined.data$LMT)

combined.data$thompson_PC3 <- (0.19*combined.data$PicVocab)+(0.067*combined.data$Flanker)+(0.493*combined.data$List)+(0.232*combined.data$CardSort)+(0.085*combined.data$Pattern)+(0.863*combined.data$Picture)+(0.122*combined.data$Reading)+(0.712*combined.data$RAVLT)+(0.068*combined.data$LMT)

# Compute a ksads "abuse" variable" and a total score
# These are the items that will go into this score (they're the physical and sexual abuse variables)
# 761, 762, 763, 767, 768, 769
combined.data$abuse <- rowSums(data.frame(as.numeric(combined.data$ksads_ptsd_raw_761_p),as.numeric(combined.data$ksads_ptsd_raw_762_p),as.numeric(combined.data$ksads_ptsd_raw_763_p),as.numeric(combined.data$ksads_ptsd_raw_767_p),as.numeric(combined.data$ksads_ptsd_raw_768_p),as.numeric(combined.data$ksads_ptsd_raw_769_p)),na.rm=TRUE)

combined.data$totalACEs <- rowSums(data.frame(as.numeric(combined.data$ksads_ptsd_raw_754_p),as.numeric(combined.data$ksads_ptsd_raw_755_p),as.numeric(combined.data$ksads_ptsd_raw_756_p),as.numeric(combined.data$ksads_ptsd_raw_757_p),as.numeric(combined.data$ksads_ptsd_raw_758_p),as.numeric(combined.data$ksads_ptsd_raw_759_p),as.numeric(combined.data$ksads_ptsd_raw_760_p),as.numeric(combined.data$ksads_ptsd_raw_761_p),as.numeric(combined.data$ksads_ptsd_raw_762_p),as.numeric(combined.data$ksads_ptsd_raw_763_p),as.numeric(combined.data$ksads_ptsd_raw_764_p),as.numeric(combined.data$ksads_ptsd_raw_765_p),as.numeric(combined.data$ksads_ptsd_raw_766_p),as.numeric(combined.data$ksads_ptsd_raw_767_p),as.numeric(combined.data$ksads_ptsd_raw_768_p),as.numeric(combined.data$ksads_ptsd_raw_769_p),as.numeric(combined.data$ksads_ptsd_raw_770_p)),na.rm=TRUE)

# Remove outliers
mean_PC1 <- mean(combined.data$thompson_PC1,na.rm=TRUE)
mean_PC2 <- mean(combined.data$thompson_PC2,na.rm=TRUE)
mean_PC3 <- mean(combined.data$thompson_PC3,na.rm=TRUE)
sd3_PC1 <- sd(combined.data$thompson_PC1,na.rm=TRUE)*3
sd3_PC2 <- sd(combined.data$thompson_PC2,na.rm=TRUE)*3
sd3_PC3 <- sd(combined.data$thompson_PC3,na.rm=TRUE)*3
sum(combined.data$thompson_PC1 > mean_PC1+sd3_PC1,na.rm=TRUE) # 8 outliers
sum(combined.data$thompson_PC2 > mean_PC2+sd3_PC2,na.rm=TRUE) # 2 outliers
sum(combined.data$thompson_PC3 > mean_PC3+sd3_PC3,na.rm=TRUE) # 0 outliers
combined.data.noOutliers <- combined.data[(combined.data$thompson_PC1>(mean_PC1-sd3_PC1) & combined.data$thompson_PC1<(mean_PC1+sd3_PC1) & combined.data$thompson_PC2>(mean_PC2-sd3_PC2) & combined.data$thompson_PC2<(mean_PC2+sd3_PC2) & combined.data$thompson_PC3>(mean_PC3-sd3_PC3) & combined.data$thompson_PC3<(mean_PC3+sd3_PC3) & combined.data$totalACEs<16 & combined.data$df18>3000),]

# calculate a total FPN score
combined.data.noOutliers$totalFPN <- rowSums(data.frame(as.numeric(combined.data.noOutliers$df4),as.numeric(combined.data.noOutliers$df16),as.numeric(combined.data.noOutliers$df18)))
```

## PFN ANALYSES
# Remember that networks 1-17 are in variables df2-18
```{r}

# Replicate findings from Cui et al., 2020
summary(lm(thompson_PC2~interview_age+sex+df18,data=combined.data.noOutliers)) #0.06
summary(lm(df18~interview_age,data=combined.data.noOutliers)) # **
summary(lm(thompson_PC2~interview_age+sex+df16,data=combined.data.noOutliers))
summary(lm(df16~interview_age,data=combined.data.noOutliers)) # **
summary(lm(thompson_PC2~interview_age+sex+df4,data=combined.data.noOutliers))
summary(lm(df4~interview_age,data=combined.data.noOutliers)) # **

# See which networks are associated with which cognitive factor scores
summary(lm(thompson_PC1~interview_age+sex+df2+df3+df4+df5+df6+df7+df8+df9+df10+df11+df12+df13+df14+df15+df16+df17+df18,data=combined.data.noOutliers))
summary(lm(thompson_PC2~interview_age+sex+df2+df3+df4+df5+df6+df7+df8+df9+df10+df11+df12+df13+df14+df15+df16+df17+df18,data=combined.data.noOutliers))
summary(lm(thompson_PC3~interview_age+sex+df2+df3+df4+df5+df6+df7+df8+df9+df10+df11+df12+df13+df14+df15+df16+df17+df18,data=combined.data.noOutliers))

summary(lm(df4~interview_age+sex+abuse,data=combined.data.noOutliers))
summary(lm(df4~interview_age+sex+totalACEs,data=combined.data.noOutliers))

summary(lm(df16~interview_age+sex+abuse,data=combined.data.noOutliers))
summary(lm(df16~interview_age+sex+totalACEs,data=combined.data.noOutliers))

summary(lm(df18~interview_age+sex+abuse,data=combined.data.noOutliers))
summary(lm(df18~interview_age+sex+totalACEs,data=combined.data.noOutliers))

summary(lm(thompson_PC1~interview_age+sex+totalACEs,data=combined.data.noOutliers)) # ***
summary(lm(thompson_PC2~interview_age+sex+totalACEs,data=combined.data.noOutliers)) # *
summary(lm(thompson_PC3~interview_age+sex+totalACEs,data=combined.data.noOutliers)) # ***
summary(lm(thompson_PC1~interview_age+sex+abuse,data=combined.data.noOutliers)) # *
summary(lm(thompson_PC2~interview_age+sex+abuse,data=combined.data.noOutliers)) #ns
summary(lm(thompson_PC3~interview_age+sex+abuse,data=combined.data.noOutliers)) # *

summary(lm(thompson_PC1~interview_age+sex+df18,data=combined.data.noOutliers)) # ***
summary(lm(thompson_PC1~interview_age+sex+df16,data=combined.data.noOutliers)) # ***
summary(lm(thompson_PC1~interview_age+sex+df4,data=combined.data.noOutliers)) # ns

# total FPN:
summary(lm(thompson_PC1~interview_age+sex+totalFPN,data=combined.data.noOutliers)) # ***
summary(lm(thompson_PC2~interview_age+sex+totalFPN,data=combined.data.noOutliers)) # .
summary(lm(thompson_PC3~interview_age+sex+totalFPN,data=combined.data.noOutliers)) # .

summary(lm(thompson_PC1~interview_age+sex+totalFPN,data=combined.data.noOutliers[combined.data.noOutliers$totalACEs==0,]))
summary(lm(thompson_PC1~interview_age+sex+totalFPN,data=combined.data.noOutliers[combined.data.noOutliers$totalACEs>0,]))

ggplot(data=combined.data.noOutliers,aes(totalFPN,thompson_PC1))+geom_point()+geom_smooth(method="lm")
ggplot(data=combined.data.noOutliers,aes(totalFPN,thompson_PC2))+geom_point()+geom_smooth(method="lm")
ggplot(data=combined.data.noOutliers,aes(totalFPN,thompson_PC3))+geom_point()+geom_smooth(method="lm")

combined.data.noOutliers$ACEsFactor <- factor(combined.data.noOutliers$totalACEs==0,labels = c("ACEs","No ACEs"))
ggplot(data=combined.data.noOutliers,aes(totalFPN,thompson_PC1,color=ACEsFactor))+geom_point(size=1)+geom_smooth(method="lm")

summary(lm(thompson_PC1~interview_age+sex+totalFPN+ACEsFactor,data=combined.data.noOutliers))
summary(lm(df18~interview_age+sex+ACEsFactor+ACEsFactor*interview_age+ACEsFactor*sex,data=combined.data.noOutliers))

ggplot(data=combined.data.noOutliers,aes(interview_age,df18,color=ACEsFactor))+geom_point(size=1)+geom_smooth(method="lm")

ggplot(data=combined.data.noOutliers,aes(interview_age,totalFPN,color=ACEsFactor))+geom_point(size=1)+geom_smooth(method="lm")
```