### 2022-06-10 ABCD_Flux22_PFN_Analyses.Rmd
analyses for flux abstract

---
title: "ABCD_Flux22_PFN_Analyses"
author: "Arielle Keller"
date: "3/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=250)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lsr)
library(lme4)
library(lmerTest)
library(mediation)
library(ggcorrplot)
library(corrplot)
library(RColorBrewer)
library(glmnet)
library(mgcv)
library(gamm4)
library(readr)

library(devtools)
install_github("jfortin1/neuroCombatData")
install_github("jfortin1/neuroCombat_Rpackage")
library(neuroCombat)
```

# First load in the spatial extent of each PFN (they're in order, 1-17)
```{r}
pfn_sizes <- read.csv("/Users/askeller/Documents/Kellernet_PrelimAnalysis/All_PFN_sizes.csv") # n=9,132
# Change PFN column names so they're sensible
colnames(pfn_sizes)<-c("subjectkey","PFN1","PFN2","PFN3","PFN4","PFN5","PFN6","PFN7","PFN8","PFN9","PFN10","PFN11","PFN12","PFN13","PFN14","PFN15","PFN16","PFN17")
```

# Remove subjects based on the following criteria:
## 1) 8min of retained TRs (600 TRs)
## 2) ABCD Booleans for rest and T1
```{r}
num_TRs <- read.csv("/Users/askeller/Documents/Kellernet_PrelimAnalysis/num_TRs_by_subj_redo.csv")
data<-merge(pfn_sizes,num_TRs,by="subjectkey")
data.clean<-data[data$numTRs>=600,] # we go from 9,132 to 8,293 at this step

# Remove participants based on booleans from ABCD
abcd_imgincl01 <- read.csv("/Users/askeller/Documents/Kellernet_PrelimAnalysis/abcd_imgincl01.csv")
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$eventnam=="baseline_year_1_arm_1",]
abcd_imgincl01 <- abcd_imgincl01[!abcd_imgincl01$visit=="",] # start here with 11,663 (8,287 of the 9,132 set)
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$imgincl_t1w_include==1,] # down to 11,124 (8,045 of the 9,132 set)
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$imgincl_rsfmri_include==1,] # down to 9,388 (7,475 of the 9,132 set)

combined.data <- merge(data.clean,abcd_imgincl01[, c("subjectkey",'imgincl_t1w_include')],by="subjectkey") # n=7,475
```

## Instead of calculating the Thompson factor scores ourselves, use the summary scores from DEAP
```{r}
newdata<-readRDS("/Users/askeller/Documents/Kellernet_PrelimAnalysis/DEAP-data-download_ThompsonScores_ResFamIncome.rds")
newdata.baseline<-newdata[newdata$event_name=="baseline_year_1_arm_1",]
colnames(newdata.baseline)[1]<-"subjectkey"
all.data <- merge(combined.data,newdata.baseline[,c("subjectkey", setdiff(colnames(newdata.baseline),colnames(combined.data)))],by="subjectkey")

# for simplicity, rename thompson variables
colnames(all.data)<-gsub("neurocog_pc1.bl","thompson_PC1",colnames(all.data))
colnames(all.data)<-gsub("neurocog_pc2.bl","thompson_PC2",colnames(all.data))
colnames(all.data)<-gsub("neurocog_pc3.bl","thompson_PC3",colnames(all.data)) # still at n=7,475
```

# Add in Family and Site covariates
```{r}
family <-read.table("/Users/askeller/Documents/Kellernet_PrelimAnalysis/ses/acspsw03.txt",header=TRUE)
family.baseline<-family[family$eventname=="baseline_year_1_arm_1",]
abcd.data.almost <- merge(all.data,family.baseline[, c("subjectkey", setdiff(colnames(family.baseline),colnames(all.data)))],by="subjectkey")
site_info <- readRDS("/Users/askeller/Documents/Kellernet_PrelimAnalysis/DEAP-siteID.rds")
site.baseline<-site_info[site_info$event_name=="baseline_year_1_arm_1",]
colnames(site.baseline)[1]<-"subjectkey"
abcd.data.almost2 <- merge(abcd.data.almost,site.baseline[,c("subjectkey", setdiff(colnames(site.baseline),colnames(abcd.data.almost)))],by="subjectkey") # n=7,474
```

# Add mean FD
```{r}
meanFD <- read.csv("/Users/askeller/Documents/Kellernet_PrelimAnalysis/PFN_Gradients_CombinedScript/getFD/meanFD_031822.csv")
colnames(meanFD)<-c("subjectkey","meanFD")
meanFD$subjectkey <- gsub(pattern="sub-NDAR",replacement="NDAR_", meanFD$subjectkey)
abcd.data <- merge(abcd.data.almost2,meanFD,by="subjectkey")

# n=7,461
```

# Load train/test split from UMinn
```{r}
traintest<-read_tsv("/Users/askeller/Documents/Kellernet_PrelimAnalysis/participants.tsv")
traintest.baseline<-traintest[traintest$session_id=="ses-baselineYear1Arm1",c("participant_id","matched_group")]
colnames(traintest.baseline)[1]<-c("subjectkey")
traintest.baseline$subjectkey <- gsub(pattern="sub-NDAR",replacement="NDAR_", traintest.baseline$subjectkey)
traintest.baseline<-traintest.baseline %>% distinct()
abcd.data.traintest <- merge(abcd.data,traintest.baseline,by="subjectkey")
abcd.data.train <- abcd.data.traintest[abcd.data.traintest$matched_group==1,] # n=3,749
abcd.data.test <- abcd.data.traintest[abcd.data.traintest$matched_group==2,] # n=3,712
```

# Apply ComBat harmonization across ABCD sites separately for train and test sets
```{r}
# prep to make sure there are no empty levels
abcd.data.train$abcd_site<-droplevels(abcd.data.train$abcd_site)
abcd.data.test$abcd_site<-droplevels(abcd.data.test$abcd_site)

# harmonize TRAIN set 
batch <- abcd.data.train$abcd_site
abcd.data.train.toharmonize <- abcd.data.train[,c("PFN1","PFN2","PFN3","PFN4","PFN5","PFN6","PFN7","PFN8","PFN9","PFN10","PFN11","PFN12","PFN13","PFN14","PFN15","PFN16","PFN17","thompson_PC1","thompson_PC2","thompson_PC3","meanFD")]
mod <- model.matrix(~abcd.data.train$interview_age+abcd.data.train$sex+abcd.data.train$rel_family_id)
dat <- t(abcd.data.train.toharmonize)
abcd.data.combat <- neuroCombat(dat=dat,mod=mod,batch=batch,eb=FALSE)
dat.harmonized<-data.frame(t(abcd.data.combat$dat.combat))
abcd.data.noPFNsCog <- subset(abcd.data.train,select=-c(PFN1,PFN2,PFN3,PFN4,PFN5,PFN6,PFN7,PFN8,PFN9,PFN10,PFN11,PFN12,PFN13,PFN14,PFN15,PFN16,PFN17,thompson_PC1,thompson_PC2,thompson_PC3,meanFD))
abcd.data.train.harmonized <- cbind(abcd.data.noPFNsCog,dat.harmonized) # Add the harmonized data back into the main dataframe (BE AWARE THAT NOT ALL DATA IS HARMONIZED)

# harmonize TEST set 
batch <- abcd.data.test$abcd_site
abcd.data.test.toharmonize <- abcd.data.test[,c("PFN1","PFN2","PFN3","PFN4","PFN5","PFN6","PFN7","PFN8","PFN9","PFN10","PFN11","PFN12","PFN13","PFN14","PFN15","PFN16","PFN17","thompson_PC1","thompson_PC2","thompson_PC3","meanFD")]
mod <- model.matrix(~abcd.data.test$interview_age+abcd.data.test$sex+abcd.data.test$rel_family_id)
dat <- t(abcd.data.test.toharmonize)
abcd.data.combat <- neuroCombat(dat=dat,mod=mod,batch=batch,eb=FALSE)
dat.harmonized<-data.frame(t(abcd.data.combat$dat.combat))
abcd.data.noPFNsCog <- subset(abcd.data.test,select=-c(PFN1,PFN2,PFN3,PFN4,PFN5,PFN6,PFN7,PFN8,PFN9,PFN10,PFN11,PFN12,PFN13,PFN14,PFN15,PFN16,PFN17,thompson_PC1,thompson_PC2,thompson_PC3,meanFD))
abcd.data.test.harmonized <- cbind(abcd.data.noPFNsCog,dat.harmonized) # Add the harmonized data back into the main dataframe (BE AWARE THAT NOT ALL DATA IS HARMONIZED)

```

# fix age variable to be continuous
```{r}
abcd.data.train.harmonized$interview_age <- as.numeric(abcd.data.train.harmonized$interview_age)
abcd.data.test.harmonized$interview_age <- as.numeric(abcd.data.test.harmonized$interview_age)
abcd.data.traintest$interview_age <- as.numeric(abcd.data.traintest$interview_age)
```

# quick  analyses for flux abstract
```{r}
summary(lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
summary(lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
summary(lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))

summary(lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
summary(lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
summary(lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
```

# Make a demographics table
```{r}
# exclude n=488 subjects with missing cognitive data
abcd.data.for.ridge<-abcd.data.traintest[,c("subjectkey","thompson_PC1","thompson_PC2","thompson_PC3","matched_group","interview_age","sex","meanFD","abcd_site","rel_family_id")]

abcd.data.traintest<-abcd.data.traintest[complete.cases(abcd.data.for.ridge),]

library(Gmisc)
demo_table <- abcd.data.traintest %>%
  mutate(matched_group = factor(matched_group, levels = 1:2, labels = c("Discovery", "Replication"))) %>%
#subset(matched_group<3) %>%
  set_column_labels(interview_age = "Age (Months)",
                    sex = "Sex",
                    race.6level = "Race",
                    household.income = "Household Income",
                    thompson_PC1 = "General Cognition",
                    thompson_PC2 = "Executive Function",
                    thompson_PC3 = "Learning/Memory",
                    high.educ = "Parent Education") %>%
getDescriptionStatsBy(interview_age, 
                        sex,
                        thompson_PC1,
                        thompson_PC2,
                        thompson_PC3,
                        household.income,
                        race.6level,
                        high.educ,
                        add_total_col = TRUE, statistics=TRUE,
                        by = matched_group)

demo_table
```

# make figures to replicate Cui et al.
```{r}
correlation<-cor.test(abcd.data.train.harmonized$thompson_PC1,abcd.data.train.harmonized$PFN17)
png("/Users/askeller/Documents/Kellernet_PrelimAnalysis/CuiRep_032322/PC1_PFN17_train.png",
    width = 3000, 
    height = 3000,
    res=1200)
ggplot(data=abcd.data.train.harmonized,aes(abcd.data.train.harmonized$thompson_PC1,abcd.data.train.harmonized$PFN17))+geom_point(size=0.1,colour="#F5BA2E")+geom_smooth(method=lm,colour="#F5BA2E")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text.x = element_text(color="black"),
    axis.ticks = element_line(color = "black"),axis.text.y = element_text(color="black")) +ylab('Total Representation') + xlab('General Cognition') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

correlation<-cor.test(abcd.data.train.harmonized$thompson_PC1,abcd.data.train.harmonized$PFN15)
png("/Users/askeller/Documents/Kellernet_PrelimAnalysis/CuiRep_032322/PC1_PFN15_train.png",
    width = 3000, 
    height = 3000,
    res=1200)
ggplot(data=abcd.data.train.harmonized,aes(abcd.data.train.harmonized$thompson_PC1,abcd.data.train.harmonized$PFN15))+geom_point(size=0.1,colour="#F5BA2E")+geom_smooth(method=lm,colour="#F5BA2E")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text.x = element_text(color="black"),
    axis.ticks = element_line(color = "black"),axis.text.y = element_text(color="black")) +ylab('Total Representation') + xlab('General Cognition') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

correlation<-cor.test(abcd.data.test.harmonized$thompson_PC1,abcd.data.test.harmonized$PFN17)
png("/Users/askeller/Documents/Kellernet_PrelimAnalysis/CuiRep_032322/PC1_PFN17_test.png",
    width = 3000, 
    height = 3000,
    res=1200)
ggplot(data=abcd.data.test.harmonized,aes(abcd.data.test.harmonized$thompson_PC1,abcd.data.test.harmonized$PFN17))+geom_point(size=0.1,colour="#F5BA2E")+geom_smooth(method=lm,colour="#F5BA2E")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text.x = element_text(color="black"),
    axis.ticks = element_line(color = "black"),axis.text.y = element_text(color="black")) +ylab('Total Representation') + xlab('General Cognition') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

correlation<-cor.test(abcd.data.test.harmonized$thompson_PC1,abcd.data.test.harmonized$PFN15)
png("/Users/askeller/Documents/Kellernet_PrelimAnalysis/CuiRep_032322/PC1_PFN15_test.png",
    width = 3000, 
    height = 3000,
    res=1200)
ggplot(data=abcd.data.test.harmonized,aes(abcd.data.test.harmonized$thompson_PC1,abcd.data.test.harmonized$PFN15))+geom_point(size=0.1,colour="#F5BA2E")+geom_smooth(method=lm,colour="#F5BA2E")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text.x = element_text(color="black"),
    axis.ticks = element_line(color = "black"),axis.text.y = element_text(color="black")) +ylab('Total Representation') + xlab('General Cognition') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

correlation<-cor.test(abcd.data.train.harmonized$thompson_PC1,abcd.data.train.harmonized$PFN3)
png("/Users/askeller/Documents/Kellernet_PrelimAnalysis/CuiRep_032322/PC1_PFN3_train.png",
    width = 3000, 
    height = 3000,
    res=1200)
ggplot(data=abcd.data.train.harmonized,aes(abcd.data.train.harmonized$thompson_PC1,abcd.data.train.harmonized$PFN3))+geom_point(size=0.1,colour="#F5BA2E")+geom_smooth(method=lm,colour="#F5BA2E")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text.x = element_text(color="black"),
    axis.ticks = element_line(color = "black"),axis.text.y = element_text(color="black")) +ylab('Total Representation') + xlab('General Cognition') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

correlation<-cor.test(abcd.data.test.harmonized$thompson_PC1,abcd.data.test.harmonized$PFN3)
png("/Users/askeller/Documents/Kellernet_PrelimAnalysis/CuiRep_032322/PC1_PFN3_test.png",
    width = 3000, 
    height = 3000,
    res=1200)
ggplot(data=abcd.data.test.harmonized,aes(abcd.data.test.harmonized$thompson_PC1,abcd.data.test.harmonized$PFN3))+geom_point(size=0.1,colour="#F5BA2E")+geom_smooth(method=lm,colour="#F5BA2E")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text.x = element_text(color="black"),
    axis.ticks = element_line(color = "black"),axis.text.y = element_text(color="black")) +ylab('Total Representation') + xlab('General Cognition') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

```

### Replicate Cui bar graph:
```{r}
pfn1<-summary(lmer(PFN1~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn2<-summary(lmer(PFN2~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn3<-summary(lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized)) 
pfn4<-summary(lmer(PFN4~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn5<-summary(lmer(PFN5~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized)) 
pfn6<-summary(lmer(PFN6~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn7<-summary(lmer(PFN7~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized)) 
pfn8<-summary(lmer(PFN8~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn9<-summary(lmer(PFN9~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn10<-summary(lmer(PFN10~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn11<-summary(lmer(PFN11~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn12<-summary(lmer(PFN12~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn13<-summary(lmer(PFN13~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn14<-summary(lmer(PFN14~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn15<-summary(lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn16<-summary(lmer(PFN16~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn17<-summary(lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))

p_vec<-c(pfn1$coefficients[5,5],pfn2$coefficients[5,5],pfn3$coefficients[5,5],pfn4$coefficients[5,5],pfn5$coefficients[5,5],pfn6$coefficients[5,5],pfn7$coefficients[5,5],pfn8$coefficients[5,5],pfn9$coefficients[5,5],pfn10$coefficients[5,5],pfn11$coefficients[5,5],pfn12$coefficients[5,5],pfn13$coefficients[5,5],pfn14$coefficients[5,5],pfn15$coefficients[5,5],pfn16$coefficients[5,5],pfn17$coefficients[5,5])

p_vec_adj <- p.adjust(p_vec, "bonferroni")
z_vec <- qnorm(p_vec/2,lower.tail = FALSE);

pos_neg<-c(pfn1$coefficients[5,1],pfn2$coefficients[5,1],pfn3$coefficients[5,1],pfn4$coefficients[5,1],pfn5$coefficients[5,1],pfn6$coefficients[5,1],pfn7$coefficients[5,1],pfn8$coefficients[5,1],pfn9$coefficients[5,1],pfn10$coefficients[5,1],pfn11$coefficients[5,1],pfn12$coefficients[5,1],pfn13$coefficients[5,1],pfn14$coefficients[5,1],pfn15$coefficients[5,1],pfn16$coefficients[5,1],pfn17$coefficients[5,1])
pos_neg <- ifelse(pos_neg>0,1,-1)

z_vec_pn <- z_vec*pos_neg
keep_p <- ifelse(p_vec_adj<.05,1,0)

```

### Replicate Cui bar graph (SAME THING IN TEST SET):
```{r}
pfn1<-summary(lmer(PFN1~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn2<-summary(lmer(PFN2~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn3<-summary(lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized)) 
pfn4<-summary(lmer(PFN4~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn5<-summary(lmer(PFN5~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized)) 
pfn6<-summary(lmer(PFN6~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn7<-summary(lmer(PFN7~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized)) 
pfn8<-summary(lmer(PFN8~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn9<-summary(lmer(PFN9~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn10<-summary(lmer(PFN10~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn11<-summary(lmer(PFN11~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn12<-summary(lmer(PFN12~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn13<-summary(lmer(PFN13~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn14<-summary(lmer(PFN14~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn15<-summary(lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn16<-summary(lmer(PFN16~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn17<-summary(lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))

p_vec<-c(pfn1$coefficients[5,5],pfn2$coefficients[5,5],pfn3$coefficients[5,5],pfn4$coefficients[5,5],pfn5$coefficients[5,5],pfn6$coefficients[5,5],pfn7$coefficients[5,5],pfn8$coefficients[5,5],pfn9$coefficients[5,5],pfn10$coefficients[5,5],pfn11$coefficients[5,5],pfn12$coefficients[5,5],pfn13$coefficients[5,5],pfn14$coefficients[5,5],pfn15$coefficients[5,5],pfn16$coefficients[5,5],pfn17$coefficients[5,5])

p_vec_adj <- p.adjust(p_vec, "bonferroni")
z_vec <- qnorm(p_vec/2,lower.tail = FALSE);

pos_neg<-c(pfn1$coefficients[5,1],pfn2$coefficients[5,1],pfn3$coefficients[5,1],pfn4$coefficients[5,1],pfn5$coefficients[5,1],pfn6$coefficients[5,1],pfn7$coefficients[5,1],pfn8$coefficients[5,1],pfn9$coefficients[5,1],pfn10$coefficients[5,1],pfn11$coefficients[5,1],pfn12$coefficients[5,1],pfn13$coefficients[5,1],pfn14$coefficients[5,1],pfn15$coefficients[5,1],pfn16$coefficients[5,1],pfn17$coefficients[5,1])
pos_neg <- ifelse(pos_neg>0,1,-1)

z_vec_pn <- z_vec*pos_neg
keep_p <- ifelse(p_vec_adj<.05,1,0)

```

## Replicate Cui result showing topography across-individual variability by network
```{r}

# library(R.matlab)
# library(ggplot2);
# 
# ResultsFolder = '/data/jux/BBL/projects/pncSingleFuncParcel/Replication/Revision';
# AtlasFolder = paste0(ResultsFolder, '/SingleParcellation/SingleAtlas_Analysis');
# Variability_Visualize_Folder = paste0(AtlasFolder, '/Variability_Visualize');
# Variability_Mat = readMat(paste0(Variability_Visualize_Folder, '/VariabilityLoading_Median_17SystemMean.mat'));
# 
# AtlasLabel_Mat = readMat(paste0(AtlasFolder, '/Group_AtlasLabel.mat'));
# 
# AllData = data.frame(Variability = matrix(0, 17734, 1));
# AllData$Label = matrix(0, 17734, 1);
# for (i in c(1:17734))
# {
#   AllData$Variability[i] = Variability_Mat$VariabilityLoading.Median.17SystemMean.NoMedialWall[i];
#   AllData$Label[i] = AtlasLabel_Mat$sbj.AtlasLabel.NoMedialWall[i];
# }
# AllData$Variability = as.numeric(AllData$Variability);
# AllData$Label = as.factor(AllData$Label);
# 
# # Order the network by the median value
# MedianValue = matrix(0, 1, 17);
# for (i in c(1:17)) {
#   ind = which(AllData$Label == i);
#   MedianValue[i] = median(AllData$Variability[ind]);
# }
# # Store network variability for correlation with effect size
# writeMat(paste0(ResultsFolder, '/Corr_NetworkVariability_EffectSize/NetworkVariability.mat'), MedianValue = MedianValue);
# tmp = sort(MedianValue, index.return = TRUE);
# 
# ColorScheme = c("#E76178", "#7499C2", "#F5BA2E", "#7499C2", "#00A131",
#                 "#AF33AD", "#E443FF", "#E76178", "#E443FF", "#AF33AD",
#                 "#7499C2", "#E76178", "#7499C2", "#00A131", "#F5BA2E",
#                 "#4E31A8", "#F5BA2E");
# ColorScheme_XText = c("#AF33AD", "#AF33AD", "#7499C2", "#7499C2", "#4E31A8",
#                 "#7499C2", "#7499C2", "#F5BA2E", "#00A131", "#E76178",
#                 "#E443FF", "#E443FF", "#E76178", "#E76178", "#F5BA2E",
#                 "#00A131", "#F5BA2E");
# Order = c(10, 6, 4, 2, 16, 11, 13, 15, 5, 12, 7, 9, 8, 1, 17, 14, 3);
# ColorScheme_XText_New = ColorScheme_XText;
# for (i in c(1:17)) {
#   ind = which(Order == i);
#   ColorScheme_XText_New[ind] = ColorScheme_XText[i];
# }
# ggplot(AllData, aes(x = Label, y = Variability, fill = Label, color = Label)) + 
#       geom_violin(trim = FALSE) +
#       scale_color_manual(values = ColorScheme) + 
#       scale_fill_manual(values = ColorScheme) + 
#       labs(x = "Networks", y = "Across-subject Variability") + theme_classic() +
#       theme(axis.text.x = element_text(size = 22, color = ColorScheme_XText_New),
#             axis.text.y = element_text(size = 22, color = "black"),
#             axis.title = element_text(size = 22)) +
#       theme(axis.text.x = element_text(angle = 22, hjust = 1)) + 
#       theme(legend.position = "none") + 
#       scale_x_discrete(
#             limits = c(10, 6, 4, 2, 16, 11, 13, 15, 5, 12, 7, 9, 8, 1, 17, 14, 3),
#             labels = c("12", "11", "2", "6", "14", "7", "8", "17", "16", "9", 
#                        "13", "5", "15", "10", "3", "1", "4")) + 
#       geom_boxplot(width=0.1, fill = "white");
# ggsave('/data/jux/BBL/projects/pncSingleFuncParcel/Replication/Revisioni/Figures/Variability_Loading_Mean_Violin.tiff', width = 20, height = 15, dpi = 600, units = "cm");

```