### 2023-04-12 Step1_TotalCorticalRepresentation/Associations_PFN_Cognition.Rmd
sensitivity analyses

---
title: "Associations_PFN_Cognition"
author: "Arielle Keller"
date: "3/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=250)
library(tidyr)
library(dplyr)
library(ggplot2)
library(lsr)
library(lme4)
library(lmerTest)
library(mediation)
library(ggcorrplot)
library(corrplot)
library(RColorBrewer)
library(glmnet)
library(mgcv)
library(gamm4)
library(readr)

library(devtools)
install_github("jfortin1/neuroCombatData")
install_github("jfortin1/neuroCombat_Rpackage")
library(neuroCombat)

library(R.matlab)
library(mgcv)
library(ggplot2)
library(visreg)
library(ggtext)
```

# First load in the spatial extent of each PFN (they're in order, 1-17)
```{r}
pfn_sizes <- read.csv("/Users/askeller/Documents/Kellernet_PrelimAnalysis/All_PFN_sizes.csv") # n=9,132
# Change PFN column names so they're sensible
colnames(pfn_sizes)<-c("subjectkey","PFN1","PFN2","PFN3","PFN4","PFN5","PFN6","PFN7","PFN8","PFN9","PFN10","PFN11","PFN12","PFN13","PFN14","PFN15","PFN16","PFN17")
pfn_sizes$subjectkey<-gsub('sub-NDARINV','NDAR_INV',pfn_sizes$subjectkey)
```

# Remove subjects based on the following criteria:
## 1) 8min of retained TRs (600 TRs)
## 2) ABCD Booleans for rest and T1
```{r}
num_TRs <- read.csv("/Users/askeller/Documents/Kellernet_PrelimAnalysis/num_TRs_by_subj_redo.csv")
data<-merge(pfn_sizes,num_TRs,by="subjectkey")
data.clean<-data[data$numTRs>=600,]

# Remove participants based on booleans from ABCD
abcd_imgincl01 <- read.csv("/Users/askeller/Documents/Kellernet_PrelimAnalysis/abcd_imgincl01.csv")
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$eventnam=="baseline_year_1_arm_1",]
abcd_imgincl01 <- abcd_imgincl01[!abcd_imgincl01$visit=="",] 
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$imgincl_t1w_include==1,]
abcd_imgincl01 <- abcd_imgincl01[abcd_imgincl01$imgincl_rsfmri_include==1,]

combined.data1 <- merge(data.clean,abcd_imgincl01[, c("subjectkey",'imgincl_t1w_include')],by="subjectkey") 
```

## Import the Thompson PC scores from DEAP
```{r}
newdata<-readRDS("/Users/askeller/Documents/Kellernet_PrelimAnalysis/DEAP-data-download_ThompsonScores_ResFamIncome.rds")
newdata.baseline<-newdata[newdata$event_name=="baseline_year_1_arm_1",]
colnames(newdata.baseline)[1]<-"subjectkey"
all.data <- merge(combined.data1,newdata.baseline[,c("subjectkey", setdiff(colnames(newdata.baseline),colnames(combined.data1)))],by="subjectkey")

# for simplicity, rename thompson variables
colnames(all.data)<-gsub("neurocog_pc1.bl","thompson_PC1",colnames(all.data))
colnames(all.data)<-gsub("neurocog_pc2.bl","thompson_PC2",colnames(all.data))
colnames(all.data)<-gsub("neurocog_pc3.bl","thompson_PC3",colnames(all.data))

```

# Add in Family and Site covariates
```{r}
family <-read.table("/Users/askeller/Documents/Kellernet_PrelimAnalysis/ses/acspsw03.txt",header=TRUE)
family.baseline<-family[family$eventname=="baseline_year_1_arm_1",]
abcd.data.almost <- merge(all.data,family.baseline[, c("subjectkey", setdiff(colnames(family.baseline),colnames(all.data)))],by="subjectkey")
site_info <- readRDS("/Users/askeller/Documents/Kellernet_PrelimAnalysis/DEAP-siteID.rds")
site.baseline<-site_info[site_info$event_name=="baseline_year_1_arm_1",]
colnames(site.baseline)[1]<-"subjectkey"
abcd.data.almost2 <- merge(abcd.data.almost,site.baseline[,c("subjectkey", setdiff(colnames(site.baseline),colnames(abcd.data.almost)))],by="subjectkey") 
```

# Add mean FD
```{r}
meanFD <- read.csv("/Users/askeller/Documents/Kellernet_PrelimAnalysis/PFN_Gradients_CombinedScript/getFD/meanFD_031822.csv")
colnames(meanFD)<-c("subjectkey","meanFD")
meanFD$subjectkey <- gsub(pattern="sub-NDAR",replacement="NDAR_", meanFD$subjectkey)
abcd.data <- merge(abcd.data.almost2,meanFD,by="subjectkey")
```

# Load discovery/replication matched samples 
```{r}
traintest<-read_tsv("/Users/askeller/Documents/Kellernet_PrelimAnalysis/participants.tsv")
traintest.baseline<-traintest[traintest$session_id=="ses-baselineYear1Arm1",c("participant_id","matched_group")]
colnames(traintest.baseline)[1]<-c("subjectkey")
traintest.baseline$subjectkey <- gsub(pattern="sub-NDAR",replacement="NDAR_", traintest.baseline$subjectkey)
traintest.baseline<-traintest.baseline %>% distinct()
abcd.data.traintest <- merge(abcd.data,traintest.baseline,by="subjectkey")
abcd.data.traintest<-abcd.data.traintest[!duplicated(abcd.data.traintest$subjectkey),] # remove duplicated subject
abcd.data.train <- abcd.data.traintest[abcd.data.traintest$matched_group==1,] 
abcd.data.test <- abcd.data.traintest[abcd.data.traintest$matched_group==2,] 
```

# Apply ComBat harmonization across ABCD sites separately for train and test sets
```{r}
# prep to make sure there are no empty levels
abcd.data.train$abcd_site<-droplevels(abcd.data.train$abcd_site)
abcd.data.test$abcd_site<-droplevels(abcd.data.test$abcd_site)

# harmonize TRAIN set 
batch <- abcd.data.train$abcd_site
abcd.data.train.toharmonize <- abcd.data.train[,c("PFN1","PFN2","PFN3","PFN4","PFN5","PFN6","PFN7","PFN8","PFN9","PFN10","PFN11","PFN12","PFN13","PFN14","PFN15","PFN16","PFN17","thompson_PC1","thompson_PC2","thompson_PC3","meanFD")]
mod <- model.matrix(~abcd.data.train$interview_age+abcd.data.train$sex+abcd.data.train$rel_family_id)
dat <- t(abcd.data.train.toharmonize)
abcd.data.combat <- neuroCombat(dat=dat,mod=mod,batch=batch,eb=FALSE)
dat.harmonized<-data.frame(t(abcd.data.combat$dat.combat))
abcd.data.noPFNsCog <- subset(abcd.data.train,select=-c(PFN1,PFN2,PFN3,PFN4,PFN5,PFN6,PFN7,PFN8,PFN9,PFN10,PFN11,PFN12,PFN13,PFN14,PFN15,PFN16,PFN17,thompson_PC1,thompson_PC2,thompson_PC3,meanFD))
abcd.data.train.harmonized <- cbind(abcd.data.noPFNsCog,dat.harmonized) # Add the harmonized data back into the main dataframe (BE AWARE THAT NOT ALL DATA IS HARMONIZED)

# harmonize TEST set 
batch <- abcd.data.test$abcd_site
abcd.data.test.toharmonize <- abcd.data.test[,c("PFN1","PFN2","PFN3","PFN4","PFN5","PFN6","PFN7","PFN8","PFN9","PFN10","PFN11","PFN12","PFN13","PFN14","PFN15","PFN16","PFN17","thompson_PC1","thompson_PC2","thompson_PC3","meanFD")]
mod <- model.matrix(~abcd.data.test$interview_age+abcd.data.test$sex+abcd.data.test$rel_family_id)
dat <- t(abcd.data.test.toharmonize)
abcd.data.combat <- neuroCombat(dat=dat,mod=mod,batch=batch,eb=FALSE)
dat.harmonized<-data.frame(t(abcd.data.combat$dat.combat))
abcd.data.noPFNsCog <- subset(abcd.data.test,select=-c(PFN1,PFN2,PFN3,PFN4,PFN5,PFN6,PFN7,PFN8,PFN9,PFN10,PFN11,PFN12,PFN13,PFN14,PFN15,PFN16,PFN17,thompson_PC1,thompson_PC2,thompson_PC3,meanFD))
abcd.data.test.harmonized <- cbind(abcd.data.noPFNsCog,dat.harmonized) # Add the harmonized data back into the main dataframe

```

# fix age variable to be continuous
```{r}
abcd.data.train.harmonized$interview_age <- as.numeric(abcd.data.train.harmonized$interview_age)
abcd.data.test.harmonized$interview_age <- as.numeric(abcd.data.test.harmonized$interview_age)
abcd.data.traintest$interview_age <- as.numeric(abcd.data.traintest$interview_age)
```

# standardize all variables for LMER (separately in Discovery and Replication samples)
```{r}
abcd.data.train.harmonized$PFN1<-(abcd.data.train.harmonized$PFN1-mean(abcd.data.train.harmonized$PFN1,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN1,na.rm=TRUE)
abcd.data.train.harmonized$PFN2<-(abcd.data.train.harmonized$PFN2-mean(abcd.data.train.harmonized$PFN2,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN2,na.rm=TRUE)
abcd.data.train.harmonized$PFN3<-(abcd.data.train.harmonized$PFN3-mean(abcd.data.train.harmonized$PFN3,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN3,na.rm=TRUE)
abcd.data.train.harmonized$PFN4<-(abcd.data.train.harmonized$PFN4-mean(abcd.data.train.harmonized$PFN4,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN4,na.rm=TRUE)
abcd.data.train.harmonized$PFN5<-(abcd.data.train.harmonized$PFN5-mean(abcd.data.train.harmonized$PFN5,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN5,na.rm=TRUE)
abcd.data.train.harmonized$PFN6<-(abcd.data.train.harmonized$PFN6-mean(abcd.data.train.harmonized$PFN6,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN6,na.rm=TRUE)
abcd.data.train.harmonized$PFN7<-(abcd.data.train.harmonized$PFN7-mean(abcd.data.train.harmonized$PFN7,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN7,na.rm=TRUE)
abcd.data.train.harmonized$PFN8<-(abcd.data.train.harmonized$PFN8-mean(abcd.data.train.harmonized$PFN8,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN8,na.rm=TRUE)
abcd.data.train.harmonized$PFN9<-(abcd.data.train.harmonized$PFN9-mean(abcd.data.train.harmonized$PFN9,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN9,na.rm=TRUE)
abcd.data.train.harmonized$PFN10<-(abcd.data.train.harmonized$PFN10-mean(abcd.data.train.harmonized$PFN10,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN10,na.rm=TRUE)
abcd.data.train.harmonized$PFN11<-(abcd.data.train.harmonized$PFN11-mean(abcd.data.train.harmonized$PFN11,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN11,na.rm=TRUE)
abcd.data.train.harmonized$PFN12<-(abcd.data.train.harmonized$PFN12-mean(abcd.data.train.harmonized$PFN12,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN12,na.rm=TRUE)
abcd.data.train.harmonized$PFN13<-(abcd.data.train.harmonized$PFN13-mean(abcd.data.train.harmonized$PFN13,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN13,na.rm=TRUE)
abcd.data.train.harmonized$PFN14<-(abcd.data.train.harmonized$PFN14-mean(abcd.data.train.harmonized$PFN14,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN14,na.rm=TRUE)
abcd.data.train.harmonized$PFN15<-(abcd.data.train.harmonized$PFN15-mean(abcd.data.train.harmonized$PFN15,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN15,na.rm=TRUE)
abcd.data.train.harmonized$PFN16<-(abcd.data.train.harmonized$PFN16-mean(abcd.data.train.harmonized$PFN16,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN16,na.rm=TRUE)
abcd.data.train.harmonized$PFN17<-(abcd.data.train.harmonized$PFN17-mean(abcd.data.train.harmonized$PFN17,na.rm=TRUE))/sd(abcd.data.train.harmonized$PFN17,na.rm=TRUE)

abcd.data.test.harmonized$PFN1<-(abcd.data.test.harmonized$PFN1-mean(abcd.data.test.harmonized$PFN1,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN1,na.rm=TRUE)
abcd.data.test.harmonized$PFN2<-(abcd.data.test.harmonized$PFN2-mean(abcd.data.test.harmonized$PFN2,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN2,na.rm=TRUE)
abcd.data.test.harmonized$PFN3<-(abcd.data.test.harmonized$PFN3-mean(abcd.data.test.harmonized$PFN3,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN3,na.rm=TRUE)
abcd.data.test.harmonized$PFN4<-(abcd.data.test.harmonized$PFN4-mean(abcd.data.test.harmonized$PFN4,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN4,na.rm=TRUE)
abcd.data.test.harmonized$PFN5<-(abcd.data.test.harmonized$PFN5-mean(abcd.data.test.harmonized$PFN5,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN5,na.rm=TRUE)
abcd.data.test.harmonized$PFN6<-(abcd.data.test.harmonized$PFN6-mean(abcd.data.test.harmonized$PFN6,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN6,na.rm=TRUE)
abcd.data.test.harmonized$PFN7<-(abcd.data.test.harmonized$PFN7-mean(abcd.data.test.harmonized$PFN7,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN7,na.rm=TRUE)
abcd.data.test.harmonized$PFN8<-(abcd.data.test.harmonized$PFN8-mean(abcd.data.test.harmonized$PFN8,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN8,na.rm=TRUE)
abcd.data.test.harmonized$PFN9<-(abcd.data.test.harmonized$PFN9-mean(abcd.data.test.harmonized$PFN9,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN9,na.rm=TRUE)
abcd.data.test.harmonized$PFN10<-(abcd.data.test.harmonized$PFN10-mean(abcd.data.test.harmonized$PFN10,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN10,na.rm=TRUE)
abcd.data.test.harmonized$PFN11<-(abcd.data.test.harmonized$PFN11-mean(abcd.data.test.harmonized$PFN11,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN11,na.rm=TRUE)
abcd.data.test.harmonized$PFN12<-(abcd.data.test.harmonized$PFN12-mean(abcd.data.test.harmonized$PFN12,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN12,na.rm=TRUE)
abcd.data.test.harmonized$PFN13<-(abcd.data.test.harmonized$PFN13-mean(abcd.data.test.harmonized$PFN13,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN13,na.rm=TRUE)
abcd.data.test.harmonized$PFN14<-(abcd.data.test.harmonized$PFN14-mean(abcd.data.test.harmonized$PFN14,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN14,na.rm=TRUE)
abcd.data.test.harmonized$PFN15<-(abcd.data.test.harmonized$PFN15-mean(abcd.data.test.harmonized$PFN15,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN15,na.rm=TRUE)
abcd.data.test.harmonized$PFN16<-(abcd.data.test.harmonized$PFN16-mean(abcd.data.test.harmonized$PFN16,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN16,na.rm=TRUE)
abcd.data.test.harmonized$PFN17<-(abcd.data.test.harmonized$PFN17-mean(abcd.data.test.harmonized$PFN17,na.rm=TRUE))/sd(abcd.data.test.harmonized$PFN17,na.rm=TRUE)

abcd.data.train.harmonized$interview_age<-(abcd.data.train.harmonized$interview_age-mean(abcd.data.train.harmonized$interview_age,na.rm=TRUE))/sd(abcd.data.train.harmonized$interview_age,na.rm=TRUE)
abcd.data.test.harmonized$interview_age<-(abcd.data.test.harmonized$interview_age-mean(abcd.data.test.harmonized$interview_age,na.rm=TRUE))/sd(abcd.data.test.harmonized$interview_age,na.rm=TRUE)

abcd.data.train.harmonized$meanFD<-(abcd.data.train.harmonized$meanFD-mean(abcd.data.train.harmonized$meanFD,na.rm=TRUE))/sd(abcd.data.train.harmonized$meanFD,na.rm=TRUE)
abcd.data.test.harmonized$meanFD<-(abcd.data.test.harmonized$meanFD-mean(abcd.data.test.harmonized$meanFD,na.rm=TRUE))/sd(abcd.data.test.harmonized$meanFD,na.rm=TRUE)

abcd.data.train.harmonized$reshist_addr1_adi_income<-(abcd.data.train.harmonized$reshist_addr1_adi_income-mean(abcd.data.train.harmonized$reshist_addr1_adi_income,na.rm=TRUE))/sd(abcd.data.train.harmonized$reshist_addr1_adi_income,na.rm=TRUE)
abcd.data.test.harmonized$reshist_addr1_adi_income<-(abcd.data.test.harmonized$reshist_addr1_adi_income-mean(abcd.data.test.harmonized$reshist_addr1_adi_income,na.rm=TRUE))/sd(abcd.data.test.harmonized$reshist_addr1_adi_income,na.rm=TRUE)
```

# linear mixed effects models associating general cognition with each fronto-parietal PFN
```{r}
summary(lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
summary(lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
summary(lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))

summary(lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
summary(lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
summary(lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))

mod1<-lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized)
mod2<-lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized)
mod3<-lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized)
mod4<-lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized)
mod5<-lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized)
mod6<-lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized)

tab_model(mod1,mod2,mod3,pred.labels = c("Intercept","Age","Sex","Mean FD","General Cognition"),dv.labels=c("PFN 3","PFN 15","PFN 17"),digits=2,digits.p=2,show.se=TRUE,show.stat=TRUE,string.stat = "t",string.se = "Std. Error",show.intercept=TRUE,show.ci = FALSE, p.adjust="none",p.style="scientific", file = "/Users/askeller/Documents/Writing_InProgress/PFNs_Cognition/Associations_TableX_Discovery.doc")

tab_model(mod4,mod5,mod6,pred.labels = c("Intercept","Age","Sex","Mean FD","General Cognition"),dv.labels=c("PFN 3","PFN 15","PFN 17"),digits=2,digits.p=2,show.se=TRUE,show.stat=TRUE,string.stat = "t",string.se = "Std. Error",show.intercept=TRUE,show.ci = FALSE, p.adjust="none",p.style="scientific", file = "/Users/askeller/Documents/Writing_InProgress/PFNs_Cognition/Associations_TableX_Replication.doc")

# sensitivity analysis with ADI as an additional covariate:
mod101<-lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+reshist_addr1_adi_income+thompson_PC1,data=abcd.data.train.harmonized)
mod102<-lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+reshist_addr1_adi_income+thompson_PC1,data=abcd.data.train.harmonized)
mod103<-lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+reshist_addr1_adi_income+thompson_PC1,data=abcd.data.train.harmonized)
mod104<-lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+reshist_addr1_adi_income+thompson_PC1,data=abcd.data.test.harmonized)
mod105<-lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+reshist_addr1_adi_income+thompson_PC1,data=abcd.data.test.harmonized)
mod106<-lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+reshist_addr1_adi_income+thompson_PC1,data=abcd.data.test.harmonized)

tab_model(mod101,mod102,mod103,pred.labels = c("Intercept","Age","Sex","Mean FD","ADI","General Cognition"),dv.labels=c("PFN 3","PFN 15","PFN 17"),digits=2,digits.p=2,show.se=TRUE,show.stat=TRUE,string.stat = "t",string.se = "Std. Error",show.intercept=TRUE,show.ci = FALSE, p.adjust="none",p.style="scientific", file = "/Users/askeller/Documents/Writing_InProgress/PFNs_Cognition/ADI_Sensitivity_TableX_Discovery.doc")

tab_model(mod104,mod105,mod106,pred.labels = c("Intercept","Age","Sex","Mean FD","ADI","General Cognition"),dv.labels=c("PFN 3","PFN 15","PFN 17"),digits=2,digits.p=2,show.se=TRUE,show.stat=TRUE,string.stat = "t",string.se = "Std. Error",show.intercept=TRUE,show.ci = FALSE, p.adjust="none",p.style="scientific", file = "/Users/askeller/Documents/Writing_InProgress/PFNs_Cognition/ADI_Sensitivity_TableX_Replication.doc")

```

# get confidence intervals
```{r}
confint(lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
confint(lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
confint(lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))

confint(lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
confint(lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
confint(lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
```

# Prep/add data for demographics table
```{r}

# exclude subjects with missing cognitive data
abcd.data.for.ridge<-abcd.data.traintest[,c("subjectkey","thompson_PC1","thompson_PC2","thompson_PC3","matched_group","interview_age","sex","meanFD","abcd_site","rel_family_id")]
abcd.data.traintest<-abcd.data.traintest[complete.cases(abcd.data.for.ridge),]

cbcl<-read.table("/Users/askeller/Documents/Kellernet_PrelimAnalysis/aseba/abcd_cbcls01.txt",header=TRUE)
cbcl_baseline <- cbcl[cbcl$eventname=="baseline_year_1_arm_1",c("subjectkey","cbcl_scr_syn_internal_r","cbcl_scr_syn_external_r","cbcl_scr_syn_totprob_r")]
cbcl_baseline$cbcl_scr_syn_internal_r <- as.numeric(cbcl_baseline$cbcl_scr_syn_internal_r)
cbcl_baseline$cbcl_scr_syn_external_r <- as.numeric(cbcl_baseline$cbcl_scr_syn_external_r)
cbcl_baseline$cbcl_scr_syn_totprob_r <- as.numeric(cbcl_baseline$cbcl_scr_syn_totprob_r)

psychMed<-read.csv("/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/PsychotropicMeds/ABCD_Psychotropic_Medications.csv")
psychMed_baseline <- psychMed[psychMed$eventname=="baseline_year_1_arm_1",c("subjectkey","ADHD_meds","Antidepressant_meds","Antipsychotic_meds")]
psychMed_baseline$ADHD_meds<-as.factor(psychMed_baseline$ADHD_meds)
psychMed_baseline$Antipsychotic_meds<-as.factor(psychMed_baseline$Antipsychotic_meds)
psychMed_baseline$Antidepressant_meds<-as.factor(psychMed_baseline$Antidepressant_meds)

abcd.data.traintest<-merge(abcd.data.traintest,cbcl_baseline,by="subjectkey")
abcd.data.traintest<-merge(abcd.data.traintest,psychMed_baseline,by="subjectkey")

abcd.data.traintest$ADHD_meds<-factor(abcd.data.traintest$ADHD_meds, levels = c("1","0"))
abcd.data.traintest$Antidepressant_meds<-factor(abcd.data.traintest$Antidepressant_meds, levels = c("1","0"))
abcd.data.traintest$Antipsychotic_meds<-factor(abcd.data.traintest$Antipsychotic_meds, levels = c("1","0"))

```

# Make a demographics table
```{r}
library(Gmisc)
demo_table <- abcd.data.traintest %>%
  mutate(matched_group = factor(matched_group, levels = 1:2, labels = c("Discovery", "Replication"))) %>%
#subset(matched_group<3) %>%
  set_column_labels(interview_age = "Age (Months)",
                    sex = "Sex (F)",
                    race.6level = "Race",
                    household.income = "Household Income",
                    high.educ = "Parent Education",
                    thompson_PC1 = "General Cognition",
                    thompson_PC2 = "Executive Function",
                    thompson_PC3 = "Learning/Memory",
                    ADHD_meds = "ADHD Medications",
                    Antidepressant_meds = "Antidepressant Medication",
                    Antipsychotic_meds = "Antipsychotic Medication",
                    cbcl_scr_syn_internal_r = "CBCL Int.",
                    cbcl_scr_syn_external_r = "CBCL Ext.",
                    cbcl_scr_syn_totprob_r = "CBCL Prob.") %>%
getDescriptionStatsBy(interview_age,
                        sex,
                        thompson_PC1,
                        thompson_PC2,
                        thompson_PC3,
                        ADHD_meds,
                        Antidepressant_meds,
                        Antipsychotic_meds,
                        cbcl_scr_syn_external_r,
                        cbcl_scr_syn_internal_r,
                        cbcl_scr_syn_totprob_r,
                        household.income,
                        race.6level,
                        high.educ,
                        add_total_col = TRUE, statistics=TRUE,
                        by = matched_group)

demo_table
```

# make figures to replicate Cui et al.
```{r}
correlation<-cor.test(abcd.data.train.harmonized$thompson_PC1,abcd.data.train.harmonized$PFN17)
png("/Users/askeller/Documents/Kellernet_PrelimAnalysis/CuiRep_032322/PC1_PFN17_train.png",
    width = 3000,
    height = 3000,
    res=1200)
ggplot(data=abcd.data.train.harmonized,aes(abcd.data.train.harmonized$thompson_PC1,abcd.data.train.harmonized$PFN17))+geom_point(size=0.1,colour="#F5BA2E")+geom_smooth(method=lm,colour="#F5BA2E")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text.x = element_text(color="black"),
    axis.ticks = element_line(color = "black"),axis.text.y = element_text(color="black")) +ylab('Total Representation') + xlab('General Cognition') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

correlation<-cor.test(abcd.data.train.harmonized$thompson_PC1,abcd.data.train.harmonized$PFN15)
png("/Users/askeller/Documents/Kellernet_PrelimAnalysis/CuiRep_032322/PC1_PFN15_train.png",
    width = 3000,
    height = 3000,
    res=1200)
ggplot(data=abcd.data.train.harmonized,aes(abcd.data.train.harmonized$thompson_PC1,abcd.data.train.harmonized$PFN15))+geom_point(size=0.1,colour="#F5BA2E")+geom_smooth(method=lm,colour="#F5BA2E")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text.x = element_text(color="black"),
    axis.ticks = element_line(color = "black"),axis.text.y = element_text(color="black")) +ylab('Total Representation') + xlab('General Cognition') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

correlation<-cor.test(abcd.data.test.harmonized$thompson_PC1,abcd.data.test.harmonized$PFN17)
png("/Users/askeller/Documents/Kellernet_PrelimAnalysis/CuiRep_032322/PC1_PFN17_test.png",
    width = 3000,
    height = 3000,
    res=1200)
ggplot(data=abcd.data.test.harmonized,aes(abcd.data.test.harmonized$thompson_PC1,abcd.data.test.harmonized$PFN17))+geom_point(size=0.1,colour="#F5BA2E")+geom_smooth(method=lm,colour="#F5BA2E")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text.x = element_text(color="black"),
    axis.ticks = element_line(color = "black"),axis.text.y = element_text(color="black")) +ylab('Total Representation') + xlab('General Cognition') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

correlation<-cor.test(abcd.data.test.harmonized$thompson_PC1,abcd.data.test.harmonized$PFN15)
png("/Users/askeller/Documents/Kellernet_PrelimAnalysis/CuiRep_032322/PC1_PFN15_test.png",
    width = 3000,
    height = 3000,
    res=1200)
ggplot(data=abcd.data.test.harmonized,aes(abcd.data.test.harmonized$thompson_PC1,abcd.data.test.harmonized$PFN15))+geom_point(size=0.1,colour="#F5BA2E")+geom_smooth(method=lm,colour="#F5BA2E")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text.x = element_text(color="black"),
    axis.ticks = element_line(color = "black"),axis.text.y = element_text(color="black")) +ylab('Total Representation') + xlab('General Cognition') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

correlation<-cor.test(abcd.data.train.harmonized$thompson_PC1,abcd.data.train.harmonized$PFN3)
png("/Users/askeller/Documents/Kellernet_PrelimAnalysis/CuiRep_032322/PC1_PFN3_train.png",
    width = 3000,
    height = 3000,
    res=1200)
ggplot(data=abcd.data.train.harmonized,aes(abcd.data.train.harmonized$thompson_PC1,abcd.data.train.harmonized$PFN3))+geom_point(size=0.1,colour="#F5BA2E")+geom_smooth(method=lm,colour="#F5BA2E")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text.x = element_text(color="black"),
    axis.ticks = element_line(color = "black"),axis.text.y = element_text(color="black")) +ylab('Total Representation') + xlab('General Cognition') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

correlation<-cor.test(abcd.data.test.harmonized$thompson_PC1,abcd.data.test.harmonized$PFN3)
png("/Users/askeller/Documents/Kellernet_PrelimAnalysis/CuiRep_032322/PC1_PFN3_test.png",
    width = 3000,
    height = 3000,
    res=1200)
ggplot(data=abcd.data.test.harmonized,aes(abcd.data.test.harmonized$thompson_PC1,abcd.data.test.harmonized$PFN3))+geom_point(size=0.1,colour="#F5BA2E")+geom_smooth(method=lm,colour="#F5BA2E")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text.x = element_text(color="black"),
    axis.ticks = element_line(color = "black"),axis.text.y = element_text(color="black")) +ylab('Total Representation') + xlab('General Cognition') + ggtitle(paste("r=",as.character(round(correlation$estimate,3)),", p=",as.character(round(correlation$p.value,3))))
dev.off()

```

# Look for effects of SES on total cortical representation
```{r}
ADI_mod1<-lmer(reshist_addr1_adi_income~interview_age+sex+(1|rel_family_id)+meanFD+PFN1+PFN2+PFN3+PFN4+PFN5+PFN6+PFN7+PFN8+PFN9+PFN10+PFN11+PFN12+PFN13+PFN14+PFN15+PFN16+PFN17,data=abcd.data.train.harmonized)
ADI_mod2<-lmer(reshist_addr1_adi_income~interview_age+sex+(1|rel_family_id)+meanFD+PFN1+PFN2+PFN3+PFN4+PFN5+PFN6+PFN7+PFN8+PFN9+PFN10+PFN11+PFN12+PFN13+PFN14+PFN15+PFN16+PFN17,data=abcd.data.test.harmonized)

tab_model(ADI_mod1,ADI_mod2,pred.labels = c("Intercept","Age","Sex","Motion","PFN1","PFN2","PFN3","PFN4","PFN5","PFN6","PFN7","PFN8","PFN9","PFN10","PFN11","PFN12","PFN13","PFN14","PFN15","PFN16","PFN17"),dv.labels=c("ADI (Discovery)","ADI (Replication)"),digits=2,digits.p=2,show.se=TRUE,show.stat=TRUE,string.stat = "t",string.se = "Std. Error",show.intercept=TRUE,show.ci = FALSE, p.adjust="bonferroni",p.style="scientific", file = "/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SES/ADI_by_PFN.doc")
```

### Bar plot (Discovery Sample):
```{r}
pfn1<-summary(lmer(PFN1~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn2<-summary(lmer(PFN2~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn3<-summary(lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized)) 
pfn4<-summary(lmer(PFN4~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn5<-summary(lmer(PFN5~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized)) 
pfn6<-summary(lmer(PFN6~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn7<-summary(lmer(PFN7~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized)) 
pfn8<-summary(lmer(PFN8~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn9<-summary(lmer(PFN9~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn10<-summary(lmer(PFN10~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn11<-summary(lmer(PFN11~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn12<-summary(lmer(PFN12~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn13<-summary(lmer(PFN13~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn14<-summary(lmer(PFN14~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn15<-summary(lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn16<-summary(lmer(PFN16~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))
pfn17<-summary(lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.train.harmonized))

p_vec<-c(pfn1$coefficients[5,5],pfn2$coefficients[5,5],pfn3$coefficients[5,5],pfn4$coefficients[5,5],pfn5$coefficients[5,5],pfn6$coefficients[5,5],pfn7$coefficients[5,5],pfn8$coefficients[5,5],pfn9$coefficients[5,5],pfn10$coefficients[5,5],pfn11$coefficients[5,5],pfn12$coefficients[5,5],pfn13$coefficients[5,5],pfn14$coefficients[5,5],pfn15$coefficients[5,5],pfn16$coefficients[5,5],pfn17$coefficients[5,5])

p_vec_adj <- p.adjust(p_vec, "bonferroni")
z_vec <- qnorm(p_vec/2,lower.tail = FALSE);

pos_neg<-c(pfn1$coefficients[5,1],pfn2$coefficients[5,1],pfn3$coefficients[5,1],pfn4$coefficients[5,1],pfn5$coefficients[5,1],pfn6$coefficients[5,1],pfn7$coefficients[5,1],pfn8$coefficients[5,1],pfn9$coefficients[5,1],pfn10$coefficients[5,1],pfn11$coefficients[5,1],pfn12$coefficients[5,1],pfn13$coefficients[5,1],pfn14$coefficients[5,1],pfn15$coefficients[5,1],pfn16$coefficients[5,1],pfn17$coefficients[5,1])
pos_neg <- ifelse(pos_neg>0,1,-1)

z_vec_pn_train <- z_vec*pos_neg
keep_p <- ifelse(p_vec_adj<.05,1,0)

z_vec_sorted = rank(z_vec_pn_train);

net_locs = c()
for (i in 1:17){
  net_locs <- c(net_locs,which(z_vec_sorted==i))
}

# Using 7 colors scheme for bar plot
# bar plot for cognition effects
data <- data.frame(CognitionEffects_Z = as.numeric(z_vec_pn_train)) 
data$EffectRank <- as.numeric(z_vec_sorted)
BorderColor <- c("#E76178", "#7499C2", "#F5BA2E", "#7499C2", "#00A131",
                 "#AF33AD", "#E443FF", "#E76178", "#E443FF", "#AF33AD",
                 "#7499C2", "#E76178", "#7499C2", "#00A131", "#F5BA2E", 
                 "#4E31A8", "#F5BA2E");
LineType <- ifelse(keep_p==1,"solid","dashed");
Fig <- ggplot(data, aes(EffectRank, CognitionEffects_Z)) +
  geom_bar(stat = "identity", fill=ifelse(keep_p==1,BorderColor,"#FFFFFF"), 
           colour = BorderColor, linetype = LineType, width = 0.8) + ylim(-8.5,5.5) +
  labs(x = "Networks", y = expression(paste("Cognition Association (", italic("Z"), ")"))) + theme_classic() +
  theme(axis.text.x = ggtext::element_markdown(size= 27, color = BorderColor[net_locs]), 
        axis.text.y = ggtext::element_markdown(size= 33, color = "black"), 
        axis.title=ggtext::element_markdown(size = 33)) +
  theme(axis.text.x = ggtext::element_markdown(angle = 45, hjust = 1)) +
  scale_x_discrete(labels = as.character(net_locs),limits = as.character(z_vec_sorted)) 
Fig
ggsave('/Users/askeller/Documents/Kellernet_PrelimAnalysis/CuiRep_032322/NetworkSize_Loading_CogEffects_Disc_Revision.tiff', width = 19, height = 16, dpi = 600, units = "cm");

```

### Bar plot (Replication Sample):
```{r}
pfn1<-summary(lmer(PFN1~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn2<-summary(lmer(PFN2~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn3<-summary(lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized)) 
pfn4<-summary(lmer(PFN4~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn5<-summary(lmer(PFN5~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized)) 
pfn6<-summary(lmer(PFN6~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn7<-summary(lmer(PFN7~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized)) 
pfn8<-summary(lmer(PFN8~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn9<-summary(lmer(PFN9~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn10<-summary(lmer(PFN10~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn11<-summary(lmer(PFN11~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn12<-summary(lmer(PFN12~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn13<-summary(lmer(PFN13~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn14<-summary(lmer(PFN14~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn15<-summary(lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn16<-summary(lmer(PFN16~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))
pfn17<-summary(lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+thompson_PC1,data=abcd.data.test.harmonized))

p_vec<-c(pfn1$coefficients[5,5],pfn2$coefficients[5,5],pfn3$coefficients[5,5],pfn4$coefficients[5,5],pfn5$coefficients[5,5],pfn6$coefficients[5,5],pfn7$coefficients[5,5],pfn8$coefficients[5,5],pfn9$coefficients[5,5],pfn10$coefficients[5,5],pfn11$coefficients[5,5],pfn12$coefficients[5,5],pfn13$coefficients[5,5],pfn14$coefficients[5,5],pfn15$coefficients[5,5],pfn16$coefficients[5,5],pfn17$coefficients[5,5])

p_vec_adj <- p.adjust(p_vec, "bonferroni")
z_vec <- qnorm(p_vec/2,lower.tail = FALSE);

pos_neg<-c(pfn1$coefficients[5,1],pfn2$coefficients[5,1],pfn3$coefficients[5,1],pfn4$coefficients[5,1],pfn5$coefficients[5,1],pfn6$coefficients[5,1],pfn7$coefficients[5,1],pfn8$coefficients[5,1],pfn9$coefficients[5,1],pfn10$coefficients[5,1],pfn11$coefficients[5,1],pfn12$coefficients[5,1],pfn13$coefficients[5,1],pfn14$coefficients[5,1],pfn15$coefficients[5,1],pfn16$coefficients[5,1],pfn17$coefficients[5,1])
pos_neg <- ifelse(pos_neg>0,1,-1)

z_vec_pn_test <- z_vec*pos_neg
keep_p <- ifelse(p_vec_adj<.05,1,0)

z_vec_sorted = rank(z_vec_pn_test);

net_locs = c()
for (i in 1:17){
  net_locs <- c(net_locs,which(z_vec_sorted==i))
}

# Using 7 colors scheme for bar plot
# bar plot for cognition effects
data <- data.frame(CognitionEffects_Z = as.numeric(z_vec_pn_test)) 
data$EffectRank <- as.numeric(z_vec_sorted)
BorderColor <- c("#E76178", "#7499C2", "#F5BA2E", "#7499C2", "#00A131",
                 "#AF33AD", "#E443FF", "#E76178", "#E443FF", "#AF33AD",
                 "#7499C2", "#E76178", "#7499C2", "#00A131", "#F5BA2E", 
                 "#4E31A8", "#F5BA2E");
LineType <- ifelse(keep_p==1,"solid","dashed");
Fig <- ggplot(data, aes(EffectRank, CognitionEffects_Z)) +
  geom_bar(stat = "identity", fill=ifelse(keep_p==1,BorderColor,"#FFFFFF"), 
           colour = BorderColor, linetype = LineType, width = 0.8) + ylim(-8.5,5.5) + 
  labs(x = "Networks", y = expression(paste("Cognition Association (", italic("Z"), ")"))) + theme_classic() +
  theme(axis.text.x = ggtext::element_markdown(size= 27, color = BorderColor[net_locs]), 
        axis.text.y = ggtext::element_markdown(size= 33, color = "black"), 
        axis.title=ggtext::element_markdown(size = 33)) +
  theme(axis.text.x = ggtext::element_markdown(angle = 45, hjust = 1)) +
  scale_x_discrete(labels = as.character(net_locs),limits = as.character(z_vec_sorted)) 
Fig
ggsave('/Users/askeller/Documents/Kellernet_PrelimAnalysis/CuiRep_032322/NetworkSize_Loading_CogEffects_Rep_Revision.tiff', width = 19, height = 16, dpi = 600, units = "cm");

```

# Test whether results hold with covariates for psychotropic med use:
```{r}
abcd.data.train.harmonized.withMeds <- merge(abcd.data.train.harmonized,abcd.data.traintest[,c("subjectkey", setdiff(colnames(abcd.data.traintest),colnames(abcd.data.train.harmonized)))],by="subjectkey")
abcd.data.test.harmonized.withMeds <- merge(abcd.data.test.harmonized,abcd.data.traintest[,c("subjectkey", setdiff(colnames(abcd.data.traintest),colnames(abcd.data.test.harmonized)))],by="subjectkey")

mod1<-lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+ADHD_meds+Antipsychotic_meds+Antidepressant_meds+thompson_PC1,data=abcd.data.train.harmonized.withMeds)
mod2<-lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+ADHD_meds+Antipsychotic_meds+Antidepressant_meds+thompson_PC1,data=abcd.data.train.harmonized.withMeds)
mod3<-lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+ADHD_meds+Antipsychotic_meds+Antidepressant_meds+thompson_PC1,data=abcd.data.train.harmonized.withMeds)
mod4<-lmer(PFN3~interview_age+sex+(1|rel_family_id)+meanFD+ADHD_meds+Antipsychotic_meds+Antidepressant_meds+thompson_PC1,data=abcd.data.test.harmonized.withMeds)
mod5<-lmer(PFN15~interview_age+sex+(1|rel_family_id)+meanFD+ADHD_meds+Antipsychotic_meds+Antidepressant_meds+thompson_PC1,data=abcd.data.test.harmonized.withMeds)
mod6<-lmer(PFN17~interview_age+sex+(1|rel_family_id)+meanFD+ADHD_meds+Antipsychotic_meds+Antidepressant_meds+thompson_PC1,data=abcd.data.test.harmonized.withMeds)

tab_model(mod1,mod2,mod3,pred.labels = c("Intercept","Age","Sex","Mean FD","ADHD Meds","Antipsychotic Meds","Antidepressant Meds","General Cognition"),dv.labels=c("PFN 3","PFN 15","PFN 17"),digits=2,digits.p=2,show.se=TRUE,show.stat=TRUE,string.stat = "t",string.se = "Std. Error",show.intercept=TRUE,show.ci = FALSE, p.adjust="none",p.style="scientific", file = "/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/PsychMedUse/Associations_TableX_Disc_PsychMed.doc")

tab_model(mod4,mod5,mod6,pred.labels = c("Intercept","Age","Sex","Mean FD","ADHD Meds","Antipsychotic Meds","Antidepressant Meds","General Cognition"),dv.labels=c("PFN 3","PFN 15","PFN 17"),digits=2,digits.p=2,show.se=TRUE,show.stat=TRUE,string.stat = "t",string.se = "Std. Error",show.intercept=TRUE,show.ci = FALSE, p.adjust="none",p.style="scientific", file = "/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/PsychMedUse/Associations_TableX_Rep_PsychMed.doc")
```