### 2023-04-12 Step6_SensitivityAnalyses/SeparateSamplesPCA/SeparateSamplesCognitionPCA.Rmd
sensitivity analyses

---
title: "SeparateSamplesCognitionPCA"
output: html_document
date: '2023-02-02'
---

# Load in libraries
```{r}
library(mvtnorm)
library(tableone)
library(parallel)
library(rstan)
library(loo)
library(gamm4)
library(Hmisc)
library(FactoMineR)
library(nFactors)
library(reshape2)
library(psych)
library(data.table)
library(mice)
library(abind)
library(cvTools)
library(modEvA)
```

# Load data:
```{r}
nihtb<-read.table("/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SeparateSamplesCognitionPCA/data/abcd_tbss01.txt",header=TRUE)
lmt<-read.table("/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SeparateSamplesCognitionPCA/data/lmtp201.txt",header=TRUE)
ravlt<-read.table("/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SeparateSamplesCognitionPCA/data/abcd_ps01.txt",header=TRUE)
site_info <- readRDS("/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SeparateSamplesCognitionPCA/data/DEAP-siteID.rds")
colnames(site_info)[1]<-"subjectkey"
matched_samples<-read_tsv("/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SeparateSamplesCognitionPCA/data/participants.tsv")
matched_samples.baseline<-matched_samples[matched_samples$session_id=="ses-baselineYear1Arm1",c("participant_id","matched_group")]
colnames(matched_samples.baseline)[1]<-c("subjectkey")
matched_samples.baseline$subjectkey <- gsub(pattern="sub-NDAR",replacement="NDAR_", matched_samples.baseline$subjectkey)
matched_samples.baseline<-matched_samples.baseline %>% distinct()
matched_samples.baseline<-matched_samples.baseline[!duplicated(matched_samples.baseline$subjectkey),] # remove duplicated subject
data_for_ridge<-read.csv("/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SeparateSamplesCognitionPCA/data/data_for_ridge_032822_subIDwithNDAR.csv")
data_for_ridge<-data_for_ridge[!duplicated(data_for_ridge$subjectkey),]
subj_to_include<-data.frame(data_for_ridge[,"subjectkey"])
colnames(subj_to_include)[1]<-"subjectkey"
```

# Convert factors to numeric values
```{r}
nihtb$nihtbx_flanker_uncorrected <- as.numeric(nihtb$nihtbx_flanker_uncorrected)
nihtb$nihtbx_picvocab_uncorrected <- as.numeric(nihtb$nihtbx_picvocab_uncorrected)
nihtb$nihtbx_pattern_uncorrected <- as.numeric(nihtb$nihtbx_pattern_uncorrected)
nihtb$nihtbx_picture_uncorrected <- as.numeric(nihtb$nihtbx_picture_uncorrected)
nihtb$nihtbx_reading_uncorrected <- as.numeric(nihtb$nihtbx_reading_uncorrected)
nihtb$nihtbx_list_uncorrected <- as.numeric(nihtb$nihtbx_list_uncorrected)
nihtb$nihtbx_cardsort_uncorrected <- as.numeric(nihtb$nihtbx_cardsort_uncorrected)
lmt$lmt_scr_perc_correct <- as.numeric(lmt$lmt_scr_perc_correct)
ravlt$pea_ravlt_sd_trial_i_tc <- as.numeric(ravlt$pea_ravlt_sd_trial_i_tc)
ravlt$pea_ravlt_sd_trial_ii_tc <- as.numeric(ravlt$pea_ravlt_sd_trial_ii_tc)
ravlt$pea_ravlt_sd_trial_iii_tc <- as.numeric(ravlt$pea_ravlt_sd_trial_iii_tc)
ravlt$pea_ravlt_sd_trial_iv_tc <- as.numeric(ravlt$pea_ravlt_sd_trial_iv_tc)
ravlt$pea_ravlt_sd_trial_v_tc <- as.numeric(ravlt$pea_ravlt_sd_trial_v_tc)
```

# Extract necessary cognition columns for just the baseline data
```{r}
nihtb_keep <- nihtb[nihtb$eventname=="baseline_year_1_arm_1",c("subjectkey","nihtbx_picvocab_uncorrected","nihtbx_flanker_uncorrected","nihtbx_pattern_uncorrected","nihtbx_reading_uncorrected","nihtbx_picture_uncorrected","nihtbx_list_uncorrected","nihtbx_cardsort_uncorrected")]
lmt_keep <- lmt[lmt$eventname=="baseline_year_1_arm_1",c("subjectkey","lmt_scr_perc_correct")]
ravlt_keep <- ravlt[ravlt$eventname=="baseline_year_1_arm_1",c("subjectkey","pea_ravlt_sd_trial_i_tc","pea_ravlt_sd_trial_ii_tc","pea_ravlt_sd_trial_iii_tc","pea_ravlt_sd_trial_iv_tc","pea_ravlt_sd_trial_v_tc")]
```

# for RAVLT we need to generate a summed score (code borrowed from Thompson et al. 2019):
```{r}
ind_pea_ravlt = c(which(names(ravlt_keep)=="pea_ravlt_sd_trial_i_tc"),which(names(ravlt_keep)=="pea_ravlt_sd_trial_ii_tc"),
                  which(names(ravlt_keep)=="pea_ravlt_sd_trial_iii_tc"),which(names(ravlt_keep)=="pea_ravlt_sd_trial_iv_tc"),
                  which(names(ravlt_keep)=="pea_ravlt_sd_trial_v_tc")); names(ravlt_keep)[ind_pea_ravlt]; summary(ravlt_keep[,ind_pea_ravlt])
ravlt_keep$RAVLT <- apply(ravlt_keep[,ind_pea_ravlt],1,sum)
```

# Combine cognitive variables into a dataframe:
```{r}
merged_cognition1<-merge(nihtb_keep,lmt_keep,by="subjectkey")
merged_cognition_final<-merge(merged_cognition1,ravlt_keep,by="subjectkey")
```

# Add in the family, site, and sample info:
```{r}
site_info_keep <- site_info[site_info$event_name=="baseline_year_1_arm_1",c("subjectkey","abcd_site","rel_family_id")]
all_sample_info <- merge(site_info_keep,matched_samples.baseline,by="subjectkey")
data_for_pca <- merge(all_sample_info,merged_cognition_final,by="subjectkey")
```

# Drop unneeded RAVLT columns, and rename all the columns
```{r}
data_for_pca_use <- data_for_pca[,c("subjectkey","abcd_site","rel_family_id","matched_group","nihtbx_picvocab_uncorrected","nihtbx_flanker_uncorrected","nihtbx_pattern_uncorrected","nihtbx_reading_uncorrected","nihtbx_picture_uncorrected","nihtbx_list_uncorrected","nihtbx_cardsort_uncorrected","lmt_scr_perc_correct","RAVLT")]
colnames(data_for_pca_use)<-c("subjectkey","Site","Family","MatchedGroup","PicVocab","Flanker","Pattern","Reading","Picture","List","CardSort","LMT","RAVLT")
data_for_pca_comp<-data_for_pca_use[complete.cases(data_for_pca_use),]
```

### IMPORTANT NOTE HERE: 3 participants who were included in the original n=6,972 do not have complete data and are therefore excluded from this analysis. Final sample will be n=6,969.

# Reduce subject list to match the imaging sample for this paper
```{r}
data_for_pca_final<-merge(data_for_pca_comp,subj_to_include,by="subjectkey")
```

# Split the sample into Discovery and Replication
```{r}
data_for_pca_Disc <- data_for_pca_final[data_for_pca_final$MatchedGroup==1,]
data_for_pca_Rep <- data_for_pca_final[data_for_pca_final$MatchedGroup==2,]
```

# Get indices for the cognitive variables, and check that they're correct
```{r}
ind_cog = c(which(names(data_for_pca_final)=="PicVocab"),which(names(data_for_pca_final)=="Flanker"),
           which(names(data_for_pca_final)=="Pattern"),which(names(data_for_pca_final)=="Picture"),
           which(names(data_for_pca_final)=="Reading"),which(names(data_for_pca_final)=="List"),
           which(names(data_for_pca_final)=="CardSort"),which(names(data_for_pca_final)=="RAVLT"),
           which(names(data_for_pca_final)=="LMT")); names(data_for_pca_final)[ind_cog]
```

# Run a simple PCA on the DISCOVERY dataset and REPLICATION dataset separately
### ASK checked the initial version of this in the full sample on 11/18/22 and verified that the top 3 tasks loading onto the top 3 PC's are the same as for the Thompson et al. 2019 factor scores
```{r}
ind_Y = ind_cog; names(data_for_pca_Disc)[ind_Y]
Y_Disc = as.matrix(scale(data_for_pca_Disc[complete.cases(data_for_pca_Disc[,c(ind_Y)]),ind_Y]))
ev_Disc = eigen(cor(Y_Disc))
ap_Disc = parallel(subject=nrow(Y_Disc),var=ncol(Y_Disc),rep=100,cent=.05)
nS_Disc = nScree(x=ev_Disc$values,aparallel=ap_Disc$eigen$qevpea)
plotnScree(nS_Disc)
ncomp = 3
y.pca.Disc = psych::principal(Y_Disc, rotate="varimax", nfactors=ncomp, scores=TRUE)
y.pca.Disc$loadings

ind_Y = ind_cog; names(data_for_pca_Rep)[ind_Y]
Y_Rep = as.matrix(scale(data_for_pca_Rep[complete.cases(data_for_pca_Rep[,c(ind_Y)]),ind_Y]))
ev_Rep = eigen(cor(Y_Rep))
ap_Rep = parallel(subject=nrow(Y_Rep),var=ncol(Y_Rep),rep=100,cent=.05)
nS_Rep = nScree(x=ev_Rep$values,aparallel=ap_Rep$eigen$qevpea)
plotnScree(nS_Rep)
ncomp = 3
y.pca.Rep = psych::principal(Y_Rep, rotate="varimax", nfactors=ncomp, scores=TRUE)
y.pca.Rep$loadings
```

# Save out the scores for each PC
```{r}
data_for_pca_Disc$SepSampPC1<-y.pca.Disc$scores[,1]
data_for_pca_Disc$SepSampPC2<-y.pca.Disc$scores[,2]
data_for_pca_Disc$SepSampPC3<-y.pca.Disc$scores[,3]

data_for_pca_Rep$SepSampPC1<-y.pca.Rep$scores[,1]
data_for_pca_Rep$SepSampPC2<-y.pca.Rep$scores[,2]
data_for_pca_Rep$SepSampPC3<-y.pca.Rep$scores[,3]

pca_input_and_results <- rbind(data_for_pca_Disc,data_for_pca_Rep)
pca_results <- pca_input_and_results[,c("subjectkey","SepSampPC1","SepSampPC2","SepSampPC3")]
data_for_ridge_6969<-merge(data_for_ridge,pca_results,by="subjectkey")
```

# Check the correlations between the Thompson scores and the separate sample PCA scores
### correlations are 0.970, 0.985 and 0.975 respectively
```{r}
cor.test(data_for_ridge_6969$thompson_PC1,data_for_ridge_6969$SepSampPC1)
cor.test(data_for_ridge_6969$thompson_PC2,data_for_ridge_6969$SepSampPC2)
cor.test(data_for_ridge_6969$thompson_PC3,data_for_ridge_6969$SepSampPC3)
```

# Create a data for ridge file that can be used to run ridge regressions using these new PC scores instead of the Thompson scores as a sensitivity analysis
```{r}
data_for_ridge_SepSampPCA<-data_for_ridge_6969[,c("subjectkey","SepSampPC1","SepSampPC2","SepSampPC3","matched_group","interview_age","sex","meanFD","abcd_site","rel_family_id")]
data_for_ridge_SepSampPCA$subjectkey<-gsub(pattern="NDAR_INV",replacement="INV", data_for_ridge_SepSampPCA$subjectkey)
write.csv(data_for_ridge_SepSampPCA,"/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SeparateSamplesCognitionPCA/data_for_ridge_SepSampPCA.csv")

```