### 2023-04-12 Step6_SensitivityAnalyses/SplitHalfReliability/splithalf_ICCs.Rmd
split half reliability

---
title: "SplitHalf_ICC"
output: html_document
date: '2023-01-25'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(psych)
library(reshape2)
library(gifti)
library(RNifti)
library(ciftiTools)
ciftiTools.setOption('wb_path', '/Users/askeller/Documents/workbench/')
library(ggplot2)

```

# Load in the csvs

```{r}
sub1_Va<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub1_Va.csv',header=FALSE)
sub1_Vb<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub1_Vb.csv',header=FALSE)
sub2_Va<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub2_Va.csv',header=FALSE)
sub2_Vb<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub2_Vb.csv',header=FALSE)
sub3_Va<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub3_Va.csv',header=FALSE)
sub3_Vb<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub3_Vb.csv',header=FALSE)
sub4_Va<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub4_Va.csv',header=FALSE)
sub4_Vb<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub4_Vb.csv',header=FALSE)
sub5_Va<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub5_Va.csv',header=FALSE)
sub5_Vb<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub5_Vb.csv',header=FALSE)
sub6_Va<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub6_Va.csv',header=FALSE)
sub6_Vb<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub6_Vb.csv',header=FALSE)
sub7_Va<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub7_Va.csv',header=FALSE)
sub7_Vb<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub7_Vb.csv',header=FALSE)
sub8_Va<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub8_Va.csv',header=FALSE)
sub8_Vb<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub8_Vb.csv',header=FALSE)
sub9_Va<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub9_Va.csv',header=FALSE)
sub9_Vb<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub9_Vb.csv',header=FALSE)
sub10_Va<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub10_Va.csv',header=FALSE)
sub10_Vb<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/SingleParcel_1by1_SH/finalV_csvs/sub10_Vb.csv',header=FALSE)
```

# Compute ICC
```{r}
ICC_results <- c()

for (sub in 1:10){
  ICCs_sub <- c()

  if (sub == 1){
    Va <- sub1_Va
    Vb <- sub1_Vb
  } else if (sub == 2){
      Va <- sub2_Va
      Vb <- sub2_Vb
  } else if (sub == 3){
      Va <- sub3_Va
      Vb <- sub3_Vb
  } else if (sub == 4){
      Va <- sub4_Va
      Vb <- sub4_Vb 
  } else if (sub == 5){
      Va <- sub5_Va
      Vb <- sub5_Vb 
  } else if (sub == 6){
      Va <- sub6_Va
      Vb <- sub6_Vb 
  } else if (sub == 7){
      Va <- sub7_Va
      Vb <- sub7_Vb 
  } else if (sub == 8){
      Va <- sub8_Va
      Vb <- sub8_Vb 
  } else if (sub == 9){
      Va <- sub9_Va
      Vb <- sub9_Vb 
  } else if (sub == 10){
      Va <- sub10_Va
      Vb <- sub10_Vb 
    }

  for (network in 1:17){
    result<-ICC(data.frame(Va[,network],Vb[,network]))$results$ICC[2]
    ICCs_sub <- c(ICCs_sub,result)
  }
  ICC_results <- rbind(ICC_results,ICCs_sub)
}

# Save results (the above step takes a while to run)
write.csv(ICC_results,'/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/ICCresults_10subj_17PFNs.csv')
```

# Just load it in
```{r}
ICC_results<-read.csv('/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/ICCresults_10subj_17PFNs.csv')
```

# Plot results

```{r}
ICC_results_df <- data.frame(ICC_results)
colnames(ICC_results_df)<-c("Default Mode (1)","Somatomotor (2)","Fronto-Parietal (3)","Somatomotor (4)","Dorsal Attention (5)","Visual (6)","Ventral Attention (7)","Default Mode (8)","Ventral Attention (9)","Visual (10)","Somatomotor (11)","Default Mode (12)","Somatomotor (13)","Dorsal Attention (14)","Fronto-Parietal (15)","Auditory (16)","Fronto-Parietal (17)")

ICC_results_melted<-melt(ICC_results_df)
ICC_results_melted$subject<-rep(1:10,17)
colnames(ICC_results_melted)<-c("PFN","ICC","Participant")

ggplot(ICC_results_melted,aes(x=ICC))+geom_histogram()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))

ggplot(ICC_results_melted,aes(x=Participant,y=ICC))+geom_point()+scale_x_discrete(limits=c(1:10))+facet_wrap(~PFN)
```
# Make workbench file to create a brainmap of average ICC by network:
```{r}
PFNs_hardparcel<-read_cifti('/Users/askeller/Documents/Kellernet_PrelimAnalysis/hardparcel_group.dscalar.nii')
map_ICC_lh <- PFNs_hardparcel$data$cortex_left
map_ICC_rh <- PFNs_hardparcel$data$cortex_right

for (PFN in 1:17) {
    map_ICC_lh[map_ICC_lh==PFN,1]=mean(ICC_results[,PFN])
    map_ICC_rh[map_ICC_rh==PFN,1]=mean(ICC_results[,PFN])
}

PFNs_hardparcel_ICC <- PFNs_hardparcel
PFNs_hardparcel_ICC$data$cortex_left <- map_ICC_lh
PFNs_hardparcel_ICC$data$cortex_right <- map_ICC_rh

write_cifti(PFNs_hardparcel_ICC,'/Users/askeller/Documents/Kellernet_PrelimAnalysis/RevisionAnalyses_PFNsCog/SplitHalfReliability/PFNs_HardParcel_ICC')
```